---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# datazoom.amazonia

<!-- badges: start -->
[![R build status](https://github.com/datazoompuc/datazoom.amazonia/workflows/R-CMD-check/badge.svg)](https://github.com/datazoompuc/datazoom.amazonia/actions)
<!-- badges: end -->

[EDIT THIS]

The goal of datazoom.amazonia is to facilitate access to official data regarding the Amazon. The package provides functions that download and pre-process selected datasets. Data is in general provided at the municipality-year level (see the documentation for more information).

**May 4th - All functions are currently under review -- please do not download till further notice!**

-   INPE - PRODES: deforestation
-   INPE - DETER: deforestation warnings
-   INPE - DEGRAD: degradation (Work In Progress)
-   MDIC - COMEX: exports and imports
-   IBGE - PIB-Munic: gdp
-   IBGE - CEMPRE: formal employment
-   IBGE - Census: income
-   IBGE - SIGMINE: mining area
-   IBGE - PAM: agricultural data
-   IBGE - PEVS: silviculture and plant extraction data
-   MAPBIOMAS: land covering
-   IMAZON - IPS: Social Progress Index by municipality
-   SEEG - Greenhouse gases emissions by municipality

## Installation

Currently you can only install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("datazoompuc/datazoom.amazonia")
```

## Usage for INPE data

```{r,eval=FALSE}
library(datazoom.amazonia)

# Downloads data

data <- load_prodes(c(2018, 2019))

data <- load_degrad(2016)

data <- load_prodes(2017, space_aggregation = "state", language = "pt")

data <- load_deter("amazonia", space_aggregation = "state",
                   time_aggregation = "year", language = "pt")

# Loads data locally

data <- load_prodes("~/Downloads")

data <- load_deter("~/Downloads")

```

## Usage for COMEX data

```{r,eval=FALSE}
# Downloads data

years <- c(2000:2009)

data <- load_comex(years, ncm = TRUE, exp = TRUE, imp = TRUE)

```

## Usage for IBGE data

```{r,eval=FALSE}
##PIB-Munic

data <- load_amazon_gdp(c(2014, 2015))

data <- load_amazon_gdp(2017, space_aggregation = "state", language = "pt")

##SIGMINE

data <- load_sigmine(space_aggregation = 'municipality')

##PAM

data <- load_pam_permanent(2013, aggregation_level = "region")

##PEVS

data <- datazoom.amazonia::load_pevs_areasilv(2017, aggregation_level = "state", language = "eng")

```

## Usage for MAPBIOMAS data

```{r,eval=FALSE}
# Downloads data

data <- load_mapbiomas_covering(space_aggregation = 'municipality', path = NULL, covering = 3, years = c(2005:2015))

data <- load_mapbiomas_transition(transition_interval = 5)

# Loads data locally

data <- load_mapbiomas_covering("~/Downloads")

data <- load_mapbiomas_transition("~/Downloads")

```

## Usage for IMAZON data

```{r,eval=FALSE}
##IPS

data <- load_IPS(download_directory = getwd(), language = "pt")

```
## Usage for SEEG data

```{r,eval=FALSE}
##SEEG

data <- load_seeg(language = "eng")

```

## Usage for SIGMINE data

```{r,eval=FALSE}
# Downloads data

data <- load_sigmine(space_aggregation = 'municipality', source = NULL, language = 'pt')

# Loads data locally

data <- load_sigmine(space_aggregation = 'state', source = "~/Downloads", language = 'eng')

```

## Credits
DataZoom is developed by a team at Pontifícia Universidade Católica do Rio de Janeiro (PUC-Rio), Department of Economics. Our official website is at: http://www.econ.puc-rio.br/datazoom.
