---
title: "CEMPRE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CEMPRE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

Employment, salary and firm count data is from IBGE's "Cadastro Central de Empresas" (aka CEMPRE). The data is available between 2006 and 2018.

The definition of the Brazilian Legal Amazon at the municipality level is taken from IBGE's "Mapas Regionais - Amazônia Legal" available here: \url{https://www.ibge.gov.br/geociencias/cartas-e-mapas/mapas-regionais/15819-amazonia-legal.html?=&t=o-que-e}. State level information for the Brazilian Legal Amazon is taken from aggregating municipalities. IMPORTANT NOTE: this means that state level information for the amazon region is not representative of the whole state, but rather represents the portion of the state that is also a part of the Brazilian Legal Amazon.

## Usage

Using the package is easy enough:

<<<<<<< HEAD
```{r setup,eval=FALSE}
# 
# library(datazoom.amazonia)
# 
# years <- c(2017:2018)
# 
# #data <- load_amazon_employment(2007)
# 
# data <- load_amazon_employment(years, space_aggregation = "state", language = "pt")
# 
# #data <- load_employment(years)
=======
```{r setup, include = FALSE}
library(datazoom.amazonia)

years <- c(2015:2018)

data <- load_amazon_employment(2007)

data <- load_amazon_employment(years, space_aggregation = "state", language = "pt")

data <- load_employment(years)
>>>>>>> master
```
