---
title: "CEMPRE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CEMPRE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview

Employment, salary and firm data from IBGE's [Cadastro Central de Empresas (CEMPRE)](https://sidra.ibge.gov.br/pesquisa/cempre/tabelas). This comprehensive dataset provides information on companies and other organizations registered with Brazil's tax authority (Receita Federal), including employment levels, wage information, and business establishment data across Brazilian municipalities and sectors.

The CEMPRE dataset is one of the most detailed sources of firm-level data in Brazil, covering virtually all formal enterprises and organizations operating in the country.

## Data Coverage

The CEMPRE dataset includes:

- **Time Period**: Annual data from 1985 to the most recent available year
- **Geographic Coverage**: All Brazilian states, regions, and municipalities
- **Sectoral Detail**: Economic activities classified by CNAE (Classificação Nacional de Atividades Econômicas)
- **Variables**: Number of establishments, total employment, average salary, payroll totals, and other firm characteristics

## Dataset Description

### Key Variables

1. **Establishment Count**: Number of registered establishments
2. **Employment**: Total number of employees and average employees per establishment
3. **Wages**: Average salary, total payroll, and wage bill information
4. **Economic Classification**: CNAE codes for sector identification
5. **Location**: State, region, and municipality identifiers

### Geographic Aggregation Levels

The data is available at three different aggregation levels:
- **Country Level**: Aggregate statistics for all of Brazil
- **State Level**: Data aggregated by state (27 units)
- **Municipality Level**: Data disaggregated to municipality level (5,570+ municipalities)

### Sectoral Detail

Data can be retrieved with sector disaggregation or aggregate form:
- **Sectoral Disaggregation**: Detailed breakdown by CNAE 2.0 (main divisions and subdivisions)
- **Aggregate**: Total across all sectors

---

## Function Parameters

**Options:**

1. **dataset**: `"cempre"`

2. **raw_data**: 
   - `TRUE`: Returns the data in its original format from IBGE
   - `FALSE`: Returns cleaned and standardized data

3. **geo_level**: 
   - `"country"`: National aggregate
   - `"state"`: Aggregated by state
   - `"municipality"`: Disaggregated to municipality level (detailed results)

4. **time_period**: Specifies the years for which data will be downloaded (e.g., `2010:2020` for 2010 through 2020)

5. **language**: 
   - `"pt"`: Portuguese language (variable names and labels)
   - `"eng"`: English language

6. **sectors**: 
   - `TRUE`: Data is returned separated and disaggregated by economic sector (CNAE)
   - `FALSE`: Data is aggregated across all sectors

--

## Practical Examples

### Example 1: National Employment Trends

```{r eval=FALSE}
library(datazoom.amazonia)

# Download treated national-level data for recent years
national_firms <- load_cempre(
  raw_data = FALSE,
  geo_level = "country",
  time_period = 2015:2022,
  language = "eng"
)

# View structure
head(national_firms)
```

### Example 2: State-Level Analysis

```{r eval=FALSE}
# Download state-level employment data
state_data <- load_cempre(
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2010:2022,
  language = "pt"
)

# Calculate average employment by state
state_summary <- state_data %>%
  group_by(state) %>%
  summarise(
    avg_employees = mean(employment, na.rm = TRUE),
    total_establishments = sum(establishments, na.rm = TRUE)
  )
```
---

## Use Cases

- **Economic Impact Analysis**: Measure employment creation and firm growth by sector and region
- **Industrial Policy Evaluation**: Assess effectiveness of development programs using firm-level data
- **Labor Market Studies**: Analyze wage trends and employment dynamics across sectors
- **Regional Development**: Compare economic structures across municipalities
- **Business Environment**: Monitor formal business registration and formalization trends

```
```

---



