---
title: "CENSO_AGRO-IBGE"
output:
  html_document:
    df_print: paged
vignette: "%\\VignetteIndexEntry{CENSO_AGRO-IBGE} %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}\n"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The census of agriculture collects information about agricultural establishments
and the agricultural activities carried out there, covering characteristics of the
producer and establishment, economy and employment in rural areas, livestock, 
farming and agroindustry. 

Data is collected by IBGE and is available at
country, state and municipality level.

------------------------------------------------------------------------

**Options:**

1.  **dataset**: `"land_area_total"`,`"area_use"`,`"employess_tractors"`,
`"land_area_producer_condition"`,`"animal_specie_production"`,
`"vegetable_production_area_type"`,`"land_area_total_mean"`,`"use_type"`,
`"employess_total_mean"`,`"tractors_total_mean"`,`"bovine_area_mean"`,
`"animal_herd_type"`,`"income_mean_vegetable_type"`,
`"vegetable_area_income_coffee_orange"`,`"production_permanent_crops"`,
`"production_temporary_crops"`,`"livestock_production"`

2.  **raw_data**: there are two options:

    - `TRUE`: if you want the data as it is originally.
    - `FALSE`: if you want the treated version of the data.

3.  **time_period**: picks the years for which the data will be
    downloaded

4.  **language**: you can choose between Portuguese `("pt")` and English
    `("eng")`
    
