---
title: "Census"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Census}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## The Dataset

There are a collection of functions that retrieve data from the census. Available census years are 1991, 2000 and 2010, although each variable may only be available in some of those years. Recently, the census employs two questionnaires, with more information being collected for a large sample of the population. Data that comes from the census sample is pointed out below.

Currently we have data on:
* Income (sample)

The definition of the Brazilian Legal Amazon at the municipality level is taken from IBGE's "Mapas Regionais - Amazônia Legal" available here: https://www.ibge.gov.br/geociencias/cartas-e-mapas/mapas-regionais/15819-amazonia-legal.html?=&t=o-que-e. State level information for the Brazilian Legal Amazon is taken from aggregating municipalities. IMPORTANT NOTE: this means that state level information for the amazon region is not representative of the whole state, but rather represents the portion of the state that is also a part of the Brazilian Legal Amazon.

## Usage

Using the package is easy enough:

```{r setup}
library(datazoom.amazonia)

data <- load_amazon_income(2010)

data <- load_amazon_income(2010, space_aggregation = "state", language = "pt")

data <- load_income(2010)
```
