---
title: "CIPÓ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CIPÓ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Mappings by [Plataforma CIPÓ](https://plataformacipo.org/) on environmental crimes. Each dataset available is a spreadsheet pertaining to a different aspect of environmental crimes, namely: the Brazilian institutions and organization involved in their prevention (the `"brazilian_actors"` dataset); the international agreements, treaties and conventions related to the subject (the `"international_cooperation"` dataset); and the arrangements regarding forest governance (the `"forest_governance"` dataset).

***
**Options:**

  1. **dataset**: there are three choices:
      * `"brazilian_actors"`: mapping of Brazilian actors involved in preventing environmental crimes.
      * `"international_cooperation"`: mapping of international agreements, treaties and conventions.
      * `"forest_governance"`: mapping of forest governance arrangements.

  2. **search**: Filters the dataset to the rows containing the chosen search parameter.
  
***

**Examples:**
 
```{r eval=FALSE}
# download the spreacdsheet on Brazilian actors involved in fighting environmental crimes
brazilian_actors <- load_cipo(dataset = "brazilian_actors")
# searching only for entries containing IBAMA
actors_ibama <- load_cipo(dataset = "brazilian_actors",
                          search = "ibama")
# entries containing IBAMA or FUNAI
actors_ibama <- load_cipo(dataset = "brazilian_actors",
                          search = "ibama|funai")
```
