---
title: "CIPÓ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CIPO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

Mappings by [Plataforma CIPÓ](https://plataformacipo.org/) on environmental crimes. Each dataset available is a spreadsheet pertaining to a different aspect of environmental crimes, namely: the Brazilian institutions and organization involved in their prevention (the `"brazilian_actors"` dataset); the international agreements, treaties and conventions related to the subject (the `"international_cooperation"` dataset); and the arrangements regarding forest governance (the `"forest_governance"` dataset).

## Usage

Using the function is easy enough:

```{r, eval=FALSE}
library(datazoom.amazonia)

# Download the full mapping on the Brazilian actors
brazilian_actors <- load_cipo(dataset = "brazilian_actors")

# Now the International cooperation mapping
cooperation <- load_cipo(dataset = "international_cooperation")

## Now taking advantage of the search parameter

# searching only for entries containing IBAMA
actors_ibama <- load_cipo(dataset = "brazilian_actors",
                          search = "ibama")

# entries containing IBAMA or FUNAI
actors_ibama <- load_cipo(dataset = "brazilian_actors",
                          search = "ibama|funai")
```
