---
title: "COMEX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{COMEX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

COMEX (Comércio Exterior - Foreign Trade) dataset provides Brazil's official international trade statistics extracted from [Siscomex](https://www.gov.br/produtividade-e-comercio-exterior/pt-br/assuntos/comercio-exterior/estatisticas/), the Integrated System of Foreign Trade maintained by the Brazilian government.

This dataset captures:

- **Export data**: Brazilian goods leaving the country, disaggregated by municipality and product
- **Import data**: Foreign goods entering Brazil, disaggregated by municipality and product  
- **Monthly frequency**: High-frequency trade data for detailed temporal analysis
- **Product classification**: Detailed product codes and descriptions
- **Geographic coverage**: Trade flows identified by Brazilian municipality
- **Long historical coverage**: Available from 1989 onwards

COMEX is the primary official source for Brazil's international trade statistics, widely used for trade policy analysis, business intelligence, academic research, and economic monitoring.

### Data Source and Coverage

COMEX data comes from:
- Official records from Siscomex (Brazil's foreign trade system)
- Mandatory declarations by exporters and importers
- Updated monthly with current month data
- Historical data from 1989 onwards

**Important note on nomenclature**: From 1989 to 1996, Brazil used a different system of product nomenclature (NBLC - Nomenclatura Brasileira de Mercadorias). All conversions to the current nomenclature system are available and the package handles this transparently.

For more information, visit the [Brazilian Ministry of Productivity, Employment and Foreign Trade](https://www.gov.br/produtividade-e-comercio-exterior/).

---

## Available Datasets

### **1. export_mun (Exports by Municipality)**

Export data disaggregated at the municipality level.

- **Coverage**: All Brazilian municipalities engaged in international trade
- **Frequency**: Monthly
- **Time period**: 1989 onwards
- **Key variables**: Export value (USD), quantity, product code, municipality, date
- **Use cases**:
  - Identify which municipalities are export hubs
  - Analyze export diversification by region
  - Track geographic shifts in export capacity
  - Municipal-level trade policy impact assessment

### **2. import_mun (Imports by Municipality)**

Import data disaggregated at the municipality level.

- **Coverage**: All Brazilian municipalities receiving imports
- **Frequency**: Monthly
- **Time period**: 1989 onwards
- **Key variables**: Import value (USD), quantity, product code, municipality, date
- **Use cases**:
  - Understand which regions import specific products
  - Analyze import dependency patterns
  - Track geographic consumption patterns
  - Regional supply chain analysis

### **3. export_prod (Exports by Producer)**

Export data organized by producer/exporter and product.

- **Coverage**: All registered exporters in Brazil
- **Frequency**: Monthly
- **Time period**: 1989 onwards
- **Key variables**: Export value (USD), quantity, product code, exporter code, date
- **Use cases**:
  - Firm-level export analysis
  - Identify major exporters and their product mix
  - Export concentration analysis
  - Exporter persistence and dynamics

### **4. import_prod (Imports by Producer)**

Import data organized by importer/distributor and product.

- **Coverage**: All registered importers in Brazil
- **Frequency**: Monthly
- **Time period**: 1989 onwards
- **Key variables**: Import value (USD), quantity, product code, importer code, date
- **Use cases**:
  - Firm-level import behavior
  - Supply chain relationships
  - Importer concentration analysis
  - International sourcing patterns

---

## Function Parameters

### 1. **dataset**

Selects which trade dataset to download.

```r
dataset = "export_mun"   # exports by municipality
dataset = "import_mun"   # imports by municipality
dataset = "export_prod"  # exports by producer/exporter
dataset = "import_prod"  # imports by producer/importer
```

### 2. **raw_data**

Controls whether to download the original data or the processed/cleaned version.

- `TRUE`: Returns raw data exactly as published by Siscomex
- `FALSE`: Returns treated data with standardized formatting, English variable names, and cleaned values

```r
raw_data = FALSE  # logical
```

### 3. **time_period**

Specifies which year(s) to download. Available from 1989 onwards.

```r
time_period = 2020              # single year
time_period = c(2018, 2020)     # specific years
time_period = 2015:2020         # range of years
```

**Note**: Monthly data means each year can be quite large. Consider downloading specific years or ranges to manage file size.

### 4. **language**

Output language for variable names and documentation.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

---
## Examples

```{r eval=FALSE}
# download treated exports data by municipality from 2020 to 2021
data <- load_br_trade(
  dataset = "export_mun",
  raw_data = FALSE,
  time_period = 2020:2021,
  language = "eng"
)

# download treated imports data by municipality from 2020 to 2021
data <- load_br_trade(
  dataset = "import_mun",
  raw_data = FALSE,
  time_period = 2020:2021,
  language = "eng"
)
```

## Data Notes

### Raw vs. Treated Data

- **Raw data** (`raw_data = TRUE`): Original Siscomex format, potentially with inconsistencies and naming conventions from different time periods
- **Treated data** (`raw_data = FALSE`): Standardized with English variable names, consistent units (USD for values), and cleaned formatting

### Product Classification

- **1989-1996**: Uses NBLC (Nomenclatura Brasileira de Mercadorias) - conversions are handled transparently
- **1997 onwards**: Uses HS (Harmonized System) classification aligned with international standards
- Product codes enable comparison with international trade databases

### Data Characteristics

1. **Monthly frequency**: Data is reported monthly; aggregation to annual or quarterly is straightforward
2. **Producer vs. Municipality**: 
   - Municipality data groups trade by geographic origin/destination
   - Producer data groups by firm/exporter-importer code
   - Use municipality for regional analysis, producer for firm analysis
3. **Missing data**: Some small trade flows may not be reported
4. **Currency**: All values in USD

### Nomenclature Conversion

When using data spanning 1989-1996 to 1997 onwards, be aware:
- Product categories may differ between nomenclature systems
- Conversions are available but not always 1:1
- Compare very old with recent data with caution

```



