---
title: "DATASUS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

DATASUS is the IT department of SUS -- the Brazilian Unified Health System. They provide data on health establishments, mortality, access to health services and several health indicators nationwide. This function allows for an easy download of several DATASUS raw datasets, and also cleans the data in a couple of datasets. The sections below explains each avaliable dataset.
  
***
**Options:**

  1. **dataset**:
      * `"datasus_sim_do"` has SIM-DO mortality data
      * Possible subsets of SIM-DO are `"datasus_sim_dofet"` (Fetal), `"datasus_sim_doext"` (External causes), `"datasus_sim_doinf"` (Children), `"datasus_sim_domat"` (Maternal)
      * `"datasus_sih_rd"` – Reduced AIHs (summary of hospitalizations)
      * `"datasus_sih_sp"` – Professional Services performed during hospitalization
      * `"datasus_sih_rj"` – Rejected AIHs (general reason)
      * `"datasus_sih_er"` – Rejected AIHs with specific error codes
      * SIH hospitalization data is split across four datasets:
      * `"datasus_cnes_lt"` has data on the number of hospital beds.
      * `"datasus_sinasc"` has information about Live Births
      * further subsets of CNES are listed later, but those only allow for the download of raw data.

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. Only effective for SIM-DO and subsets, SIH, and CNES-LT. 
      
  3. **keep_all**: only applies when raw_data is FALSE. There are two options:
      * `TRUE`: keeps all original variables, adding variable labels and possibly constructing extra variables.
      * `FALSE`: aggregates data at the municipality, thereby losing individual-level data, and only keeping aggregate measures.
  
  4. **time_period**: picks the years for which the data will be downloaded
  
  5. **states**: a vector of states by which to filter the data. Only works for datasets whose data is provided in separate files by state.

  6. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
***

##### DATASUS - SIM (System of Mortality Information)
Each original SIM data file contains rows corresponding to a declaration of death (DO), and columns with several characteristics of the person, the place of death, and the cause of death. The data comes from the main SIM-DO (Declarations of Death) dataset, which goes by the option `"datasus_sim_do"`. There are also 4 subsets of SIM-DO, namely SIM-DOFET (Fetal), SIM-DOMAT (Maternal), SIM-DOINF (Children), and SIM-DOEXT (External Causes), with corresponding dataset options `"datasus_sim_dofet", "datasus_sim_domat", "datasus_sim_doinf", "datasus_sim_doext"`. Note that only SIM-DO provides separate files for each state, so all other dataset options always contain data from the whole country.

Below is an example of downloading the raw data, and also using the `raw_data = FALSE` option to obtain treated data. When this option is selected, we create several variables for deaths from each cause, which are encoded by their CID-10 codes. The function then returns, by default, the aggregated data of mortality sources at the municipality level. In this process, all the individual information such as age, sex, race, and schooling are lost, so we also offer the option of `keep_all = TRUE`, which creates all the indicator variables for cause of death, adds variable labels, and does not aggregate, thereby keeping all individual-level variables.

**Examples:**

```{r, eval = FALSE}
library(datazoom.amazonia)

# download raw data for the year 2010 in the state of AM.
data <- load_datasus(
  dataset = "datasus_sim_do",
  time_period = 2010,
  states = "AM",
  raw_data = TRUE
)

# download treated data with the number of deaths by cause in AM and PA.
data <- load_datasus(
  dataset = "datasus_sim_do",
  time_period = 2010,
  states = c("AM", "PA"),
  raw_data = FALSE
)

# download treated data with the number of deaths by cause in AM and PA
# keeping all individual variables.
data <- load_datasus(
  dataset = "datasus_sim_do",
  time_period = 2010,
  states = c("AM", "PA"),
  raw_data = FALSE,
  keep_all = TRUE
)
```

##### DATASUS - CNES (National Register of Health Establishments)
Provides information on health establishments, avaliable hospital beds, and active physicians. The data is split into 13 datasets: LT (Beds), ST (Establishments), DC (Complimentary data), EQ (Equipment), SR (Specialized services), HB (License), PF (Practitioner), EP (Teams), RC (Contractual Rules), IN (Incentives), EE (Teaching establishments), EF (Philanthropic establishments), and GM (Management and goals).

Raw data is avaliable for all of them using the dataset option `datasus_cnes_lt, datasus_cnes_st`, and so on, and treated data is only avaliable for CNES - LT. When `raw_data = FALSE` is chosen, we return data on the number of total hospital beds and the ones avaliable through SUS, which can be aggregated by municipality (with option `keep_all = FALSE`) or keeping all original variables (`keep_all = TRUE`).

**Examples:**

```{r, eval = FALSE}
library(datazoom.amazonia)

# download treated data with the number of avaliable beds in AM and PA.
data <- load_datasus(
  dataset = "datasus_cnes_lt",
  time_period = 2010,
  states = c("AM", "PA"),
  raw_data = FALSE
)
```

##### DATASUS – SIH (Hospital Information System)
SIH consists of multiple datasets that record detailed information about hospital admissions funded by Brazil's public health system (SUS). Each row corresponds to a Hospital Admission Authorization (AIH), and the files are organized by the type of information they contain. The main available datasets are:

`"datasus_sih_rd"` – Reduced AIH (AIH Reduzida)
Contains consolidated information about approved and processed AIHs, including the main procedure performed, related diagnoses, and total costs. This is the most commonly used dataset for statistical and epidemiological analyses.

`"datasus_sih_sp" ` – Professional Services (Serviços Profissionais)
Provides detailed records of the professional services carried out during hospital stays, including procedures performed, professionals involved (CBO/CNS), and amounts paid for medical and hospital services.

`"datasus_sih_rj"` – Rejected AIHs (AIHs Rejeitadas)
Includes consolidated records of AIHs that were rejected, specifying the general reason for the rejection but without detailed error codes. Useful for analyzing the volume and impact of rejections.

`"datasus_sih_er"` – Rejected AIHs with Error Codes (AIHs com Erro)
Contains AIHs that were rejected due to inconsistencies found during processing. Each rejection includes a specific error code indicating the reason (e.g., invalid patient data, procedure incompatibilities).

All SIH datasets are available by state and time period. By using the `raw_data = TRUE` parameter, the original DATASUS files are downloaded in their segmented form. If `raw_data = FALSE` is used, the data is returned as a single, processed tibble with renamed variables and bilingual (Portuguese/English) variable descriptions, making analysis easier and more intuitive.

**Examples:**

```{r, eval = FALSE}
library(datazoom.amazonia)

# Download of raw data for Reduced AIH (AIH Reduzida) – State of Amazonas, 2010
data_rd_raw <- load_datasus(
  dataset = "datasus_sih_rd",
  time_period = 2010,
  states = "AM",
  raw_data = TRUE
)

# Download of processed data for Rejected AIHs with Error Codes – State of Amazonas, 2010
data_er_processed <- load_datasus(
  dataset = "datasus_sih_er",
  time_period = 2010,
  states = "AM",
  raw_data = FALSE
)

# Download of raw data for Professional Services – State of Acre, 2010
data_sp_raw <- load_datasus(
  dataset = "datasus_sih_sp",
  time_period = 2010,
  states = "AC",
  raw_data = TRUE
)

# Download of processed data for Professional Services – Federal District, 2010
data_sp_processed <- load_datasus(
  dataset = "datasus_sih_sp",
  time_period = 2010,
  states = "DF",
  raw_data = FALSE
)
```

##### DATASUS – SINASC (Live Birth Information System)
Datasus Sinasc is a dataset maintained by the Live Birth Information System (SINASC), which collects and records detailed information about births in Brazil. The data is extracted from Live Birth Certificates (DNVs) and includes information about the newborn, such as sex, weight, gestational age, as well as data about the mother, such as age, number of children and health conditions. SINASC is essential for monitoring maternal and child health and generating relevant indicators for the formulation of public health policies.

The raw data is available through the `raw_data = TRUE` parameter in the `datasus_sinasc` dataset option. If the `raw_data = FALSE` parameter is chosen, the processed data will be provided, with detailed columns, renamed variables and description of the variables in both Portuguese and English, facilitating the analysis and interpretation of the data.

**Examples:**

```{r, eval = FALSE}
library(datazoom.amazonia)

# download raw data
data <- load_datasus(
  dataset = "datasus_sinasc",
  time_period = 2015,
  states = "AM",
  raw_data = TRUE
)

# download data in a single tibble, with variable labels
data <- load_datasus(
  dataset = "datasus_sinasc",
  time_period = 2015,
  states = "AM",
  raw_data = FALSE
)
```
