---
title: "DEGRAD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DEGRAD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [DEGRAD project](http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/degrad) uses satellites to monitor degradation of forest areas. Raw data is available as simple features (sf) objects, read from shapefiles. The project was substituted in 2016 by DETER-B. Accordingly, data is available from 2007 up to 2016.

Original documentation for this data is very scarce, users beware. Some things to keep in mind are:

Event data is organized through yearly editions (DEGRAD 2007-2016). Inside a given edition however, there may be data from different years (events that happened in 2015 inside DEGRAD 2016 for example). 

This package provides degradation data with municipality identification. It does this by intersecting DEGRAD geometries with IBGE's municipality geometries from the year 2019. CRS metadata however is missing from the original data source. A best effort approach is used and a CRS is assumed `(proj4string: "+proj=longlat +ellps=aust_SA +towgs84=-66.8700,4.3700,-38.5200,0.0,0.0,0.0,0.0 +no_defs")`.

***
**Options:**

  1. **dataset**: `"degrad"`

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. 
  
  3. **time_period**: picks the years for which the data will be downloaded

  4. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
***

**Examples:**

```{r, eval = FALSE}
# download treated data (raw_data = TRUE) related to forest degradation
# from 2010 to 2012 (time_period = 2010:2012). 
data <- load_degrad(dataset = 'degrad', 
                    raw_data = FALSE,
                    time_period = 2010:2012)
```
