---
title: "DEGRAD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DEGRAD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This dataset in currently under development.

## The Dataset
The DEGRAD project uses sattelites to monitor degradation of forest areas. Raw data is available as simple features (sf) objects, read from shapefiles. The project was substituted in 2016 by DETER-B. Accordingly, data is available from 2007 up to 2016.

Original documentation for this data is very scarce, users beware. Some things to keep in mind are:

Event data is organized through yearly editions (DEGRAD 2007-2016). Inside a given edition however, there may be data from different years (events that happened in 2015 inside DEGRAD 2016 for example). 

<!-- This package provides degradation data by municipality-year (or municipality-month, when available). It does this by intersecting DEGRAD geometries with IBGE's municipality geometries from the year 2019. CRS metadata however is missing from the original data source. A best effort approach is used and a CRS is assumed (proj4string: "+proj=longlat +ellps=aust_SA +towgs84=-66.8700,4.3700,-38.5200,0.0,0.0,0.0,0.0 +no_defs"). -->

## Usage

Using the package is easy enough:

```{r, eval = FALSE}
library(datazoom.amazonia)

# download treated data from 2007 to 2016
degrad_all <- load_degrad(dataset = "degrad", raw_data = FALSE, time_period = 2007:2016)

# download raw data from 2007 to 2016
raw_degrad_all <- load_degrad(dataset = "degrad", raw_data = TRUE, time_period = 2007:2016)
```
