---
title: "DEGRAD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DEGRAD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The [DEGRAD project](http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/degrad) is a research initiative that uses satellite imagery to monitor forest degradation in the Amazon. Unlike DETER's near real-time alerts, DEGRAD provides a more detailed annual analysis of forest degradation patterns.

This dataset captures:

- **Forest degradation monitoring**: Tracks areas where forests are being damaged without complete clearing
- **Annual editions**: Data released as yearly reports with accumulated observations
- **Spatial polygons**: Detailed geographic boundaries of degradation events
- **Municipality linkage**: Enhanced version links degradation areas to affected municipalities
- **Historical coverage**: Multiple years of degradation monitoring from 2007 onwards

DEGRAD data is valuable for understanding forest degradation as a distinct phenomenon from clear-cut deforestation, important for carbon accounting, biodiversity protection, and understanding transition stages toward complete forest loss.

### Data Source and Methodology

DEGRAD monitoring:
- Conducted by INPE's forest monitoring programs
- Uses satellite imagery interpretation to identify forest degradation signs
- Focuses on selective logging, small-scale agriculture, forest fires, and other degrading activities
- Released as annual editions with comprehensive analysis
- Limited documentation available (original INPE documentation is sparse)

For information, visit [INPE Forest Monitoring](http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/degrad).

***

## Important Data Characteristics

### Data Organization

 **Important**: DEGRAD data is organized differently than real-time systems. Key points:

1. **Yearly editions**: Data is organized by publication year (e.g., "DEGRAD 2016"), not event year
2. **Mixed event years**: A DEGRAD edition may contain degradation events from different years
   - Example: DEGRAD 2016 edition may include events detected in 2015 or even earlier
3. **Documentation limited**: Original INPE documentation is minimal; users should be aware of potential inconsistencies

### Spatial Integration

This package enhances the raw DEGRAD data by:
- Intersecting DEGRAD spatial polygons with IBGE municipality boundaries (2019 version)
- Providing municipality identification for each degradation event
- Converting to Simple Features (SF) objects for spatial analysis

**Note on CRS**: Coordinate system metadata should be verified after loading, as original INPE data sometimes has unclear CRS information.

***

## Available Dataset

### **degrad (Forest Degradation)**
Detailed monitoring of forest degradation across the Legal Amazon.

- **Coverage**: Legal Amazon region with focus on degradation detection
- **Time period**: 2007 onwards (but events within editions may vary)
- **Spatial unit**: Polygons/spatial geometries with municipality identification
- **Variables**: Event date/year, degradation type/cause, area, municipality, state, edition
- **Data format**: Simple Features (SF) spatial objects with geographic boundaries
- **Use cases**:
  - Distinguish degradation from complete deforestation
  - Analyze forest degradation hotspots
  - Understand selective logging extent
  - Carbon stock assessment
  - Forest fire impacts
  - Long-term degradation trends

***

## Function Parameters

### 1. **dataset**

Only one dataset is available:

```r
dataset = "degrad"  # Forest degradation monitoring
```

### 2. **raw_data**

Controls whether to download the original data or the processed/enhanced version.

- `TRUE`: Returns raw INPE data with minimal processing
- `FALSE`: Returns treated data with English variable names, municipality identification from spatial intersection, and standardized formatting

```r
raw_data = FALSE  # logical
```

**Recommendation**: Use `raw_data = FALSE` for most applications to get municipality-level information.
### 3. **time_period**

Specifies which year(s) of degradation events to download.

- **Available range**: Generally 2007-2016 (check current availability)
- **Format**: Single year, vector of years, or range

**Important**: When you request a year, you get events from that year regardless of which DEGRAD edition they appear in.

```r
time_period = 2015              # single year
time_period = c(2010, 2015)     # multiple specific years
time_period = 2010:2015         # range of years
```

### 4. **language**

Output language for variable names and documentation.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

***

## Data Structure

The returned data is a Simple Features (SF) spatial object with:

- **Spatial column**: Geometric polygons representing degradation areas
- **year**: Year when degradation was detected/event occurred
- **degradation_type**: Type of degradation (logging, fire, agriculture, etc.)
- **area_hectares**: Size of degradation event
- **municipality**: Name of affected municipality
- **state**: Brazilian state
- **edition**: Which DEGRAD edition this event appears in
- **Additional attributes**: Quality metrics, confidence levels (vary by edition)

***


## Examples

```{r eval=FALSE}
# download treated forest degradation data from 2010 to 2012
data <- load_degrad(
  dataset = "degrad",
  raw_data = FALSE,
  time_period = 2010:2012,
  language = "eng"
)
```

## Data Notes

### Data Organization Complexity

The annual edition structure (e.g., "DEGRAD 2016") mixed with variable event years within those editions means:
- When you request year 2015, you get all detected 2015 events regardless of edition
- Some 2015 events may appear in both DEGRAD 2015 and DEGRAD 2016 editions
- Duplication is handled in the data loading process

### Degradation Types

Common degradation types include:
- **Selective logging**: Commercial timber extraction
- **Forest fires**: Fire damage to forest areas
- **Agricultural clearing**: Small-scale farming expansion
- **Mining**: Degradation from mining activities
- **Other**: Mixed or unclassified degradation causes

(Exact categories vary by edition; verify with your loaded data)

### Spatial Considerations

1. **Polygons not points**: Each event is a geometric polygon, not a single location point
2. **Municipality intersection**: Treated data identifies all municipalities polygon overlaps
3. **CRS verification**: Check coordinate system after loading
4. **Geometry validity**: Some polygons may have validity issues; use `st_is_valid()` to check

### Data Limitations

1. **Limited documentation**: INPE's original documentation for DEGRAD is sparse
2. **Mixed time periods**: Events from different years appear in same edition
3. **Possible inconsistencies**: Classification and methodology may vary across editions
4. **Detection limits**: Minimum detectable degradation size varies by methodology/edition
5. **Not real-time**: This is annual analysis, not near-real-time detection like DETER

***
