---
title: "DETER"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DETER}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

DETER uses satellite surveillance to detect and report changes in forest cover across the Legal Amazon and the Cerrado biome. Each data point consists of a warning, describing which type of change has affected a certain area of forest at a given date. Broadly speaking, it makes a distinction between events of deforestation, degradation and logging. The data extracted here spans from 2016 onward in the Amazon, and from 2018 onward in the Cerrado. 

## Usage

```{r,eval=FALSE}
library(datazoom.amazonia)

# Download raw data (raw_data = TRUE) from Amazonia (dataset = "deter_amz")
deter_amz <- load_deter(dataset = 'deter_amz',
                        raw_data = TRUE)

# Download treated data (raw_data = FALSE) from Cerrado (dataset = "deter_cerrado")
# in portuguese (language = 'pt')
deter_cer <- load_deter(dataset = 'deter_cerrado',
                        raw_data = FALSE,
                        language = "pt")
```

