---
title: "DETER"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DETER}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

[DETER](http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/deter/deter) (Real-Time Detection System) is a satellite-based monitoring system operated by [INPE](https://www.inpe.br/) (National Institute of Space Research) that detects and reports changes in forest cover with near real-time frequency.
This system provides:

- **Forest cover change detection**: Identifies deforestation and forest degradation events
- **Near real-time monitoring**: Updates frequently (typically daily or weekly)
- **Spatial precision**: Geolocated detection of disturbances with geographic coordinates
- **Biome coverage**: Monitors Legal Amazon and Cerrado biome regions
- **Event-based data**: Each detection is a separate record with location and date
- **Historical records**: Accumulated database of past detection events

DETER is the primary early-warning system for deforestation in the Amazon, used by Brazilian environmental agencies for enforcement, research institutions for analysis, and internationally for monitoring compliance with forest conservation goals.

### Data Source and Methodology

DETER monitoring:
- Uses satellite imagery from multiple sources (MODIS, Landsat, Sentinel)
- Automated and manual analysis to detect recent disturbances
- Generates "alerts" - polyons representing areas of detected change
- Updates available regularly (frequency varies by satellite availability)
- Operated by INPE's Remote Sensing Division

For technical details, visit [INPE DETER System](http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/deter/deter).

---

## Available Datasets

### **1. deter_amz (DETER Amazon)**

DETER monitoring data for the Legal Amazon biome.

- **Geographic coverage**: Legal Amazon region (approximately 5 million km²)
- **Biome**: Tropical forest
- **Time period**: Historical data spanning multiple years with continuing updates
- **Spatial unit**: Polyons/spatial geometries with municipality identification
- **Variables**: Detection date, deforestation/degradation type, area, municipality, state
- **Update frequency**: Regular updates (typically daily or weekly)
- **Use cases**:
  - Monitor recent deforestation hotspots
  - Analyze deforestation trends in specific regions
  - Track enforcement impact on forest cover
  - Early-warning system for forest loss
  - Academic research on deforestation drivers

### **2. deter_cerrado (DETER Cerrado)**

DETER monitoring data for the Cerrado biome.

- **Geographic coverage**: Cerrado region (Brazilian tropical savanna)
- **Biome**: Tropical savanna/grassland
- **Time period**: Historical data with continuing updates
- **Spatial unit**: Polygons/spatial geometries with municipality identification
- **Variables**: Detection date, vegetation loss type, area, municipality, state
- **Update frequency**: Regular updates
- **Use cases**:
  - Monitor Cerrado vegetation loss (increasingly threatened biome)
  - Analyze agricultural expansion in Cerrado region
  - Compare forest and savanna degradation patterns
  - Track conservation effectiveness

---

## Important Data Characteristics

### Raw Data Structure

The raw DETER data from INPE reports one alert/detection per row, with each row typically associated with a single municipality in the raw data. However, **many alerts actually overlap multiple municipalities** (typically 2-4 municipalities) that are not all shown in the original records.
### Data Processing

This package provides an important enhancement: **spatial intersection with IBGE municipality geometries** (2019 version) to identify ALL municipalities that each DETER alert overlaps with. This creates a more complete and accurate geographic picture than the original raw data.

**Important note on CRS metadata**: The CRS (Coordinate Reference System) information may need verification after loading, as coordinate system metadata can sometimes be unclear in the original INPE data.

---

## Function Parameters

### 1. **dataset**

Selects which biome's DETER data to download.

```r
dataset = "deter_amz"      # Legal Amazon monitoring
dataset = "deter_cerrado"  # Cerrado biome monitoring
```

### 2. **raw_data**

Controls whether to download the original data or the processed/enhanced version.

- `TRUE`: Returns raw INPE data with limited municipality information
- `FALSE`: Returns treated data with spatial intersection to identify all affected municipalities, standardized English variable names

```r
raw_data = FALSE  # logical
```

**Recommendation**: Use `raw_data = FALSE` to get the enhanced municipality identification from spatial intersection.

### 3. **language**

Output language for variable names and documentation.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```


---

## Examples

```{r eval=FALSE}
# download treated DETER Amazon data
deter_amz <- load_deter(
  dataset = "deter_amz",
  raw_data = FALSE,
  language = "eng"
)

# download treated DETER Cerrado data
deter_cerrado <- load_deter(
  dataset = "deter_cerrado",
  raw_data = FALSE,
  language = "eng"
)
```

## Data Notes

### Raw Data Limitations

- **Original INPE data**: Shows one alert per municipality, even when alerts overlap multiple municipalities
- **Enhanced version**: Spatial intersection with IBGE 2019 municipality boundaries identifies all affected municipalities
- **Advantage of treated data**: More complete geographic picture of deforestation extent

### Data Structure

Each alert/row typically contains:
- **Spatial geometry**: Polygon coordinates (SF object)
- **Detection date**: When the alert was issued
- **Alert type**: Deforestation, forest degradation, or other disturbance
- **Area**: Size of detected change in hectares
- **Municipality**: Geographic unit identification
- **State**: Brazilian state
- **Metadata**: Satellite source, confidence level (varies by product)

### Important Considerations

1. **Alert types vary**: "Deforestation" is permanent forest loss; "degradation" is forest damage but not clear-cut
2. **Near real-time data**: Alerts are issued frequently; data is continuously updated
3. **Minimum detection size**: Varies by sensor; typically 25 hectares for Amazon, larger for Cerrado
4. **CRS metadata**: Verify coordinate system after loading; typically UTM zones for Brazil
5. **Overlapping municipalities**: Enhanced version accounts for alerts crossing municipality boundaries
6. **False positives possible**: Satellite detection can occasionally misclassify cloud shadows or other features as forest loss
