---
title: "DETER"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DETER}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

[DETER](http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/deter/deter) (Real-Time Detection System) is a satellite-based monitoring system operated by [INPE](https://www.inpe.br/) (National Institute of Space Research) that detects and reports changes in forest cover with near real-time frequency.
This system provides:

- **Forest cover change detection**: Identifies deforestation and forest degradation events
- **Near real-time monitoring**: Updates frequently (typically daily or weekly)
- **Spatial precision**: Geolocated detection of disturbances with geographic coordinates
- **Biome coverage**: Monitors the Legal Amazon, the Cerrado and Pantanal biomes, and non-forest areas of the Legal Amazon
- **Event-based data**: Each detection is a separate record with location and date
- **Historical records**: Accumulated database of past detection events

DETER is the primary early-warning system for deforestation in the Amazon, used by Brazilian environmental agencies for enforcement, research institutions for analysis, and internationally for monitoring compliance with forest conservation goals.

### Data Source and Methodology

DETER monitoring:
- Uses imagery from wide-field optical satellites (the satellites listed in the raw data are CBERS-4, CBERS-4A and Amazonia-1 and, for the Amazon product only, Resourcesat-2)
- Automated and manual analysis to detect recent disturbances
- Generates "alerts" - polyons representing areas of detected change
- Updates available regularly (frequency varies by satellite availability)
- Operated by INPE's Remote Sensing Division

For technical details, visit [INPE DETER System](http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/deter/deter).

***

## Available Datasets

### **1. deter_amz (DETER Amazon)**

DETER monitoring data for the Legal Amazon biome.

- **Geographic coverage**: Legal Amazon region (approximately 5 million km²)
- **Biome**: Tropical forest
- **Time period**: Alerts from August 2016 onward, with continuing updates
- **Spatial unit**: Polyons/spatial geometries with municipality identification
- **Variables**: Detection date, deforestation/degradation type, area, municipality, state
- **Alert types** (`alert_type`): `DESMATAMENTO_CR`, `DESMATAMENTO_VEG`, `DEGRADACAO`, `CICATRIZ_DE_QUEIMADA`, `CORTE_SELETIVO`, `CS_DESORDENADO`, `CS_GEOMETRICO` and `MINERACAO`
- **Update frequency**: Regular updates (typically daily or weekly)
- **Use cases**:
  - Monitor recent deforestation hotspots
  - Analyze deforestation trends in specific regions
  - Track enforcement impact on forest cover
  - Early-warning system for forest loss
  - Academic research on deforestation drivers

### **2. deter_cerrado (DETER Cerrado)**

DETER monitoring data for the Cerrado biome.

- **Geographic coverage**: Cerrado region (Brazilian tropical savanna)
- **Biome**: Tropical savanna/grassland
- **Time period**: Alerts from May 2018 onward, with continuing updates
- **Spatial unit**: Polygons/spatial geometries with municipality identification
- **Variables**: Detection date, vegetation loss type, area, municipality, state
- **Alert types** (`alert_type`): only `DESMATAMENTO_CR` (clear-cut deforestation)
- **Update frequency**: Regular updates
- **Use cases**:
  - Monitor Cerrado vegetation loss (increasingly threatened biome)
  - Analyze agricultural expansion in Cerrado region
  - Compare forest and savanna degradation patterns
  - Track conservation effectiveness

### **3. deter_pantanal (DETER Pantanal)**

DETER monitoring data for the Pantanal biome.

- **Geographic coverage**: Pantanal region (the world's largest tropical wetland), in the states of Mato Grosso do Sul and Mato Grosso
- **Biome**: Tropical wetland
- **Time period**: Alerts from August 2023 onward, with continuing updates
- **Spatial unit**: Polygons/spatial geometries with municipality identification
- **Variables**: Detection date, alert type, area, municipality, state
- **Alert types** (`alert_type`): lowercase Portuguese labels -- `cicatriz de queimada` (burn scar), `mineracao` (mining), `supressao com solo exposto` (suppression with exposed soil) and `supressao com vegetacao` (suppression with vegetation)
- **Update frequency**: Regular updates
- **Use cases**:
  - Monitor fire scars and vegetation removal in the Pantanal
  - Track pressure on the wetland alongside the Cerrado and Amazon products

### **4. deter_non_forest (DETER Non-Forest, Legal Amazon)**

DETER monitoring data for the non-forest areas of the Legal Amazon, published by INPE as a product separate from `deter_amz`.

- **Geographic coverage**: Non-forest areas within the Legal Amazon states (AC, AM, AP, MA, MT, PA, RO, RR, TO)
- **Biome**: Non-forest areas of the Legal Amazon
- **Time period**: Alerts from August 2023 onward, with continuing updates
- **Spatial unit**: Polygons/spatial geometries with municipality identification
- **Variables**: Detection date, alert type, area, municipality, state
- **Alert types** (`alert_type`): the same lowercase Portuguese labels as `deter_pantanal`
- **Update frequency**: Regular updates
- **Use cases**:
  - Monitor vegetation loss in the Legal Amazon's non-forest areas
  - Complement the Amazon forest alerts (`deter_amz`) with this separate product

**Note on class vocabularies**: the raw source data uses different class labels across these products -- upper-snake-case codes for `deter_amz` and `deter_cerrado`, lowercase Portuguese labels for `deter_pantanal` and `deter_non_forest`. `load_deter()` does not normalize them: `alert_type` keeps whichever vocabulary the source used, so recode it yourself before stacking datasets from different products.

***

## Important Data Characteristics

### Raw Data Structure

The raw DETER data from INPE reports one alert/detection per row, with each row typically associated with a single municipality in the raw data. However, **many alerts actually overlap multiple municipalities** (typically 2-4 municipalities) that are not all shown in the original records.
### Data Processing

This package provides an important enhancement: **spatial intersection with IBGE municipality geometries** (2019 version) to identify ALL municipalities that each DETER alert overlaps with. This creates a more complete and accurate geographic picture than the original raw data.

**Important note on CRS metadata**: The CRS (Coordinate Reference System) information may need verification after loading, as coordinate system metadata can sometimes be unclear in the original INPE data.

***

## Function Parameters

### 1. **dataset**

Selects which biome's DETER data to download.

```r
dataset = "deter_amz"         # Legal Amazon monitoring
dataset = "deter_cerrado"     # Cerrado biome monitoring
dataset = "deter_pantanal"    # Pantanal biome monitoring
dataset = "deter_non_forest"  # Non-forest areas of the Legal Amazon
```

### 2. **raw_data**

Controls whether to download the original data or the processed/enhanced version.

- `TRUE`: Returns raw INPE data with limited municipality information
- `FALSE`: Returns treated data with spatial intersection to identify all affected municipalities, standardized English variable names

```r
raw_data = FALSE  # logical
```

**Recommendation**: Use `raw_data = FALSE` to get the enhanced municipality identification from spatial intersection.

### 3. **language**

Output language for variable names and documentation.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```


***

## Examples

```{r eval=FALSE}
# download treated DETER Amazon data
deter_amz <- load_deter(
  dataset = "deter_amz",
  raw_data = FALSE,
  language = "eng"
)

# download treated DETER Cerrado data
deter_cerrado <- load_deter(
  dataset = "deter_cerrado",
  raw_data = FALSE,
  language = "eng"
)

# download treated DETER Pantanal data
deter_pantanal <- load_deter(
  dataset = "deter_pantanal",
  raw_data = FALSE,
  language = "eng"
)

# download treated DETER non-forest (Legal Amazon) data
deter_non_forest <- load_deter(
  dataset = "deter_non_forest",
  raw_data = FALSE,
  language = "eng"
)
```

## Data Notes

### Raw Data Limitations

- **Original INPE data**: Shows one alert per municipality, even when alerts overlap multiple municipalities
- **Enhanced version**: Spatial intersection with IBGE 2019 municipality boundaries identifies all affected municipalities
- **Advantage of treated data**: More complete geographic picture of deforestation extent

### Data Structure

Each alert/row typically contains:
- **Spatial geometry**: Polygon coordinates (SF object)
- **Detection date**: When the alert was issued
- **Alert type**: Deforestation, forest degradation, or other disturbance (the label vocabulary differs between products, see above)
- **Area**: Size of detected change in square metres (a `units` object, `[m^2]`)
- **Municipality**: Geographic unit identification
- **State**: Brazilian state
- **Metadata**: Satellite source, confidence level (varies by product)

### Important Considerations

1. **Alert types vary**: "Deforestation" is permanent forest loss; "degradation" is forest damage but not clear-cut
2. **Near real-time data**: Alerts are issued frequently; data is continuously updated
3. **Minimum detection size**: Varies by sensor; typically 25 hectares for Amazon, larger for Cerrado
4. **CRS metadata**: `load_deter()` reprojects internally to IBGE's official equal-area projection for area calculations (Albers, SIRGAS 2000/GRS80, central meridian -54) before computing `area` and overlaying municipality boundaries -- not UTM zones. See `R/deter.R`'s `# FRAGILE:` comment above `operation_crs` for the exact parameters and source.
5. **Overlapping municipalities**: Enhanced version accounts for alerts crossing municipality boundaries
6. **False positives possible**: Satellite detection can occasionally misclassify cloud shadows or other features as forest loss

***
