---
title: "DETER"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DETER}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

DETER uses satellite surveillance to detect and report changes in forest cover across the Legal Amazon and the Cerrado biome. Each data point consists of a warning, describing which type of change has affected a certain area of forest at a given date. Broadly speaking, it makes a distinction between events of deforestation, degradation and logging. 

Areas are reported in square kilometers and informed as both the area which intersects a Conservation Unit and the area intersecting a Municipality. The data extracted here spans from 2016 onward in the Amazon, and from 2018 onward in the Cerrado. 

## Usage

```{r,eval=FALSE}
library(datazoom.amazonia)

# download raw data from all years from Amazonia
deter_amz <- load_deter(dataset = 'deter_amz',
                        raw_data = TRUE,
                        time_period = "all")
```

