---
title: "DETER"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DETER}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

DETER uses satellite surveillance to detect and report changes in forest cover across the Legal Amazon and the Cerrado biome. Each data point consists of a warning, describing which type of change has affected a certain area of forest at a given date. Broadly speaking, it makes a distinction between events of deforestation, degradation and logging. 

Areas are reported in square kilometers and informed as both the area which intersects a Conservation Unit and the area intersecting a Municipality. The data extracted here spans from 2016 onward in the Amazon, and from 2018 onward in the Cerrado. 

## Usage

```{r setup}
library(datazoom.amazonia)

load_deter("amazonia", space_aggregation = "municipality", time_aggregation = "year", language = "eng")
```

Note that data extracted via 'load_deter_raw()' retains some additional information on the satellites used on each observation, as well as a variable saying which Conservation Unit was affected in each warning, if any.
