---
title: "EPE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EPE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Loads data from Energy Research Company (EPE), a brazilian public company that works closely with the brazilian Ministry of Mines and Energy (MME) and other agencies to assure the sustainable development of Brazil's energy infraestructure. EPE´s duty on that mission is to support MME with research in order to aid Brazil's energy infraestructure planning.

As of now, there are two different datasets available for download.

The energy consumption per class dataset provides data about monthly energy consumption and consumers from 2004 to 2022, whith three different agregattion levels. 
The Region level encompasses the five brazillian geographical regions (North, Northeast, Midwest, Southeast and South).
The Subsystem level encompasses the five brazillian Electric Subsystems (North, Northeast, Southeast, South/Midwest, Isolated Systems)
The State level encompasses 


[descrição do BEN]

***
**Options:**

  1. **dataset**: there are two choices:
      * `"energy_consumption_per_class"`: energy consumption and consumers by State or Subsystem
      * `"national_energy_balance"`:

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. 
      
  3. **geo_level**: only applies to the `"energy_consumption_per_class"` dataset. 
      * `"state"`
      * `"subsystem"`

  4. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
***

**Examples:**  
  
```{r eval=FALSE}
# download treated data about energy consumption at the state level
clean_epe <- load_epe(
  dataset = "energy_consumption_per_class",
  geo_level = "state",
  raw_data = FALSE
)
```
