---
title: "GDP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GDP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

GDP data at current prices is from IBGE's "Produto Interno Bruto dos Municípios" (aka PIB-Munic) and Population data is from IBGE's "Estimativas de População". GDP per capita data is simply obtained by taking the ratio of these two columns. The data is available between 2002 and 2017.

The definition of the Brazilian Legal Amazon at the municipality level is taken from IBGE's "Mapas Regionais - Amazônia Legal" available here: https://www.ibge.gov.br/geociencias/cartas-e-mapas/mapas-regionais/15819-amazonia-legal.html?=&t=o-que-e. State level information for the Brazilian Legal Amazon is taken from aggregating municipalities. IMPORTANT NOTE: this means that state level information for the amazon region is not representative of the whole state, but rather represents the portion of the state that is also a part of the Brazilian Legal Amazon.

## Usage

Using the package is easy enough:

```{r,eval=FALSE}
library(datazoom.amazonia)

years <- c(2015:2017)

data <- load_amazon_gdp(years)

data <- load_amazon_gdp(2017, space_aggregation = "state", language = "pt")

data <- load_gdp(years)
```
