---
title: "GOOGLEDRIVE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GOOGLEDRIVE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For some of our functions, the original data is stored in Google Drive and exceeds the file size limit for which direct downloads are possible. As a result, the `googledrive` package is required to download the data though the Google Drive API and run the function.

The first time the package is called, it requires you to link your Google account and grant permissions to be able to download data through the Google Drive API.

You **must** tick all boxes when the permissions page opens, or else the following error will occur:

```{r}
#Error in `gargle_abort_request_failed()`:
#! Client error: (403) Forbidden
#Insufficient Permission: Request had insufficient authentication scopes.
#• domain: global
#• reason: insufficientPermissions
#• message: Insufficient Permission: Request had insufficient authentication
#  scopes.
#Run `rlang::last_error()` to see where the error occurred.
```

For further information, click [here](https://googledrive.tidyverse.org/) to access the official package page.
