---
title: "IBAMA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IBAMA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

The dataset is originally from the Brazilian Institute of Environment and Renewable  Natural Resources (Ibama), documenting environmental embargoes and fines at the individual level from 2005 to the present day.

## Usage

`load_ibama` downloads and compiles data on environmental fines at the municipality level considering the Amazon region. 

The function returns either the raw data or a data frame with aggregates considering, for eachtime-location period, counts for total the number of infractions, infractions that already went to trial, and number of unique perpetrators of infractions.

```{r, eval = FALSE}
library(datazoom.amazonia)

# download raw data from all country
raw_ibama_all <- load_ibama(dataset = "areas_embargadas",
                            raw_data = TRUE,
                            legal_amazon_only = FALSE)
```
