---
title: "IEMA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IEMA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

The dataset is originally from the Institute of Environment and Water Resources (IEMA), documenting the number of people without access to eletric energy throughout the Amazon region in the year 2018.

## Usage

`load_iema` downloads and compiles data on the population without access to eletric energy at the municipality level, considering the Amazon region. 

The function returns only the treated data considering municipality, number of the population and the state.

```{r, eval = FALSE}
library(datazoom.amazonia)

# Download treated data (raw_data = FALSE)
data <- load_iema(dataset = "iema", raw_data = FALSE,
                     geo_level = "municipality", language = "pt")

# Download treated data in english
data <- load_iema(dataset = "iema", raw_data = FALSE,
                     geo_level = "municipality", language = "eng")
```
