---
title: "IMAZON"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IMAZON}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview

[Imazon](https://imazon.org.br/) (Instituto do Homem e Meio Ambiente da Amazônia - Institute of Man and Environment of the Amazon) is an independent Brazilian research organization that produces its own deforestation monitoring and analysis. 

This dataset provides:

- **Municipality-level deforestation pressure classification**: Risk assessment of deforestation threat
- **Pressure categories**: Three-level classification system (0, 1, 2, 3) representing deforestation pressure intensity
- **Geographic coverage**: Municipalities in the Legal Amazon region
- **Municipal boundaries**: Provided as spatial features (shapefile format)
- **Policy-relevant indicator**: Used in environmental monitoring and policy assessments

Imazon's deforestation pressure index helps identify which municipalities face high deforestation risk, valuable for targeted conservation, enforcement, and development planning.

### Data Source

Imazon's classification is based on:
- Deforestation monitoring using satellite imagery
- Analysis of deforestation trends and drivers
- Risk assessment methodology to classify municipalities
- Regular updates as monitoring data accumulates

For more information, visit [Imazon Official Website](https://imazon.org.br/).

***

## Available Dataset

### **imazon_shp (Imazon Municipality Pressure Index)**

Municipality-level deforestation pressure classification with spatial geometries.

- **Geographic unit**: Brazilian municipalities
- **Coverage**: Legal Amazon municipalities
- **Variables**: Municipality name, state, deforestation pressure category (0-3)
- **Data format**: Shapefile / Simple Features (SF) spatial objects
- **Update frequency**: Periodic updates as Imazon releases new assessments
- **Use cases**:
  - Identify high-risk deforestation municipalities
  - Target conservation resources to threatened areas
  - Analyze geographic patterns of deforestation pressure
  - Environmental risk assessment
  - Monitoring enforcement effectiveness
  - Correlate pressure with socioeconomic variables

***

## Deforestation Pressure Categories

Imazon's three-level classification system:
| Category | Pressure Level | Description |
|----------|---|---|
| 0 | No/Low Pressure | Minimal deforestation threat |
| 1 | Moderate Pressure | Some deforestation risk |
| 2 | High Pressure | Significant deforestation threat |
| 3 | Very High Pressure | Severe/Critical deforestation risk |

***

## Function Parameters

### 1. **dataset**

Only one dataset is available:

```r
dataset = "imazon_shp"  # Imazon municipality pressure classification
```

### 2. **raw_data**

Controls whether to download the original data or the processed/cleaned version.

- `TRUE`: Returns raw shapefile data exactly as provided by Imazon
- `FALSE`: Returns treated data with standardized English variable names, consistent formatting, and SF object structure

```r
raw_data = FALSE  # logical
```

### 3. **language**

Output language for variable names and documentation.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

***

## Examples

```{r eval=FALSE}
# download treated Imazon deforestation pressure data
data <- load_imazon(
  raw_data = FALSE,
  language = "eng"
)
```

## Data Notes

### Data Format

- **Shapefile format**: Returned as Simple Features (SF) spatial objects
- **Geometric representation**: Municipality polygons with pressure classification
- **Attributes**: Municipality name, state, pressure category

### Pressure Categories

The 0-3 scale represents Imazon's assessment of deforestation risk:
- Based on historical deforestation trends
- Current forest cover
- Accessibility and economic drivers
- Human pressure factors

### Important Limitations

1. **Static classification**: This is a classification/index, not time-series deforestation data
2. **Methodology may change**: Imazon's methodology for calculating pressure may be updated
3. **Subjective risk assessment**: Classification involves judgment in addition to objective metrics
4. **Not current alerts**: This is not real-time detection like DETER; it's a classification product
5. **Municipal-level only**: Finer spatial resolution not available

### Using with Other Datasets

This dataset works well in combination with:
- **DETER**: Compare pressure classification with actual recent deforestation
- **DEGRAD**: Correlate pressure with degradation patterns
- **CEMPRE**: Analyze relationship between employment and deforestation pressure
- **COMEX**: Examine trade patterns in high-pressure vs low-pressure areas

***
