---
title: "IMAZON"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IEMA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

Loads data containing the municipality's categorization according to imazon's categories. There are four categories that correspond to the level of deforestation pressure faced by the place.

## Usage

`load_imazon` downloads and compiles data that categorizes the level of deforestation of all the municipalities from the legal amazon region.

The function only offers the raw_data, as imazon already delivers the organized data.

```{r, eval = FALSE}
library(datazoom.amazonia)

# Download treated data (raw_data = FALSE)
data <- load_imazon(dataset = "imazon_shp", raw_data = FALSE,
                     geo_level = "municipality", language = "pt")

# Download treated data in english
data <- load_iema(dataset = "iema", raw_data = FALSE,
                     geo_level = "municipality", language = "eng")
```
