---
title: "Social Progress Index"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Social Progress Index}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The [Amazon Social Progress Index (IPS)](https://imazon.org.br/) is a comprehensive indicator framework that measures social and environmental progress in the Legal Amazon region. This collaborative initiative combines:

- **Imazon** (Instituto do Homem e Meio Ambiente da Amazônia): Brazilian research organization
- **Social Progress Imperative**: International organization focused on measuring societal well-being

This dataset captures:

- **Multi-dimensional development indicators**: Spanning 8 domains of social and environmental progress
- **Municipality-level data**: All Legal Amazon municipalities assessed
- **Quality of life metrics**: Health, education, sanitation, infrastructure
- **Environmental indicators**: Forest cover, deforestation risk, sustainability
- **Violence and safety**: Public safety and security metrics
- **Temporal coverage**: Data from 2014, 2018, 2021, 2023
- **Geographic coverage**: 570+ municipalities across Legal Amazon

The IPS provides a holistic view of sustainable development, moving beyond simple economic measures (GDP) to encompass environmental sustainability and social well-being.

### Data Source and Methodology

The Social Progress Index:
- Based on 50+ individual indicators across 12 domains
- Uses data from government agencies, NGOs, and research institutions
- Aggregated into 3 main dimensions and 12 subdimensions
- Indexed to 0-100 scale for comparability
- Methodologically rigorous with transparent weighting

For detailed methodology, visit [Social Progress Imperative](https://www.socialprogress.org/).

***

## Available Dimensions

The IPS framework includes 8 main dataset options:

### **1. all**
Complete Social Progress Index with all dimensions and indicators.

- **Coverage**: Comprehensive assessment across all domains
- **Variables**: All indicators and index scores
- **Use cases**: Holistic development analysis, overall progress tracking, multi-dimensional comparisons

### **2. life_quality**
Indicators related to quality of life and well-being.

- **Variables**: Healthcare quality, life expectancy, nutrition, shelter quality
- **Use cases**: Health and wellness analysis, living standards assessment, healthcare quality evaluation

### **3. sanit_habit**
Sanitation and habitat indicators.

- **Variables**: Access to improved sanitation, water quality, housing conditions
- **Use cases**: Infrastructure assessment, water and sanitation access analysis, housing quality evaluation

### **4. violence**
Public safety and violence indicators.

- **Variables**: Crime rates, safety perceptions, homicide data
- **Use cases**: Public safety analysis, violence hotspot identification, security trends

### **5. educ**
Education and literacy indicators.

- **Variables**: School enrollment, literacy rates, educational attainment, quality of education
- **Use cases**: Education access analysis, literacy trends, human capital assessment

### **6. communic**
Communication and connectivity indicators.

- **Variables**: Internet access, mobile phone coverage, communication infrastructure
- **Use cases**: Digital divide analysis, connectivity assessment, tech adoption patterns

### **7. mortality**
Health and mortality indicators.

- **Variables**: Child mortality, maternal mortality, mortality rates by cause
- **Use cases**: Health outcomes analysis, maternal/child health assessment, disease burden evaluation

### **8. deforest**
Environmental and deforestation indicators.

- **Variables**: Forest cover, deforestation rates, environmental sustainability
- **Use cases**: Forest monitoring, environmental assessment, climate/conservation analysis

***

## Function Parameters

### 1. **dataset**

Selects which dimension(s) to download.

```r
dataset = "all"         # All dimensions
dataset = "life_quality" # Quality of life metrics
dataset = "sanit_habit"  # Sanitation and habitat
dataset = "violence"     # Public safety and violence
dataset = "educ"         # Education indicators
dataset = "communic"     # Communication and connectivity
dataset = "mortality"    # Health and mortality
dataset = "deforest"     # Environmental and deforestation
```

### 2. **raw_data**

Controls whether to download original or processed data.

- `TRUE`: Returns raw data exactly as published
- `FALSE`: Returns treated data with standardized English variable names and formatting

```r
raw_data = FALSE  # logical
```

### 3. **time_period**

Specifies which assessment year(s) to download.

**Available years**: 2014, 2018, 2021, 2023

```r
time_period = 2023              # Most recent
time_period = c(2018, 2023)     # Specific years
time_period = c(2014, 2018, 2021, 2023)  # Multiple years
```

### 4. **language**

Output language for variable names and labels.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

***

## Examples

### Example 1: Overall Social Progress Assessment

Get a comprehensive view of development progress:

```{r eval=FALSE}
library(dplyr)

# Load complete IPS data for most recent year
ips_complete <- load_ips(
  dataset = "all",
  raw_data = FALSE,
  time_period = 2023,
  language = "eng"
)

# Top municipalities by overall progress
top_performers <- ips_complete %>%
  arrange(desc(overall_index)) %>%
  select(municipality, state, overall_index) %>%
  head(20)

print(top_performers)

# Summary statistics
summary(ips_complete$overall_index)
```

### Example 2: Education Access and Quality

Analyze educational indicators across the Amazon:

```{r eval=FALSE}
library(dplyr)

# Load education dimension
education_data <- load_ips(
  dataset = "educ",
  raw_data = FALSE,
  time_period = 2023,
  language = "eng"
)

# States with highest education index
state_education <- education_data %>%
  group_by(state) %>%
  summarize(
    avg_education_index = mean(education_index, na.rm = TRUE),
    num_municipalities = n(),
    min_index = min(education_index, na.rm = TRUE),
    max_index = max(education_index, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(avg_education_index))

print(state_education)

# Municipalities with lowest education access (priority areas)
priority_education <- education_data %>%
  arrange(education_index) %>%
  select(municipality, state, education_index) %>%
  head(30)

print(priority_education)
```
***

## Data Notes

### Index Scales

- **0-100 scale**: All indices standardized to 0-100 for comparison
- **Higher is better**: Across all dimensions except deforestation (where higher forest index = better)
- **Comparable across dimensions**: Standardized scale allows cross-dimension comparison

### Dimensions and Indicators

Each dimension contains multiple indicators:
- **Life quality**: 4-6 indicators
- **Sanitation/habitat**: 3-5 indicators  
- **Violence**: 3-4 indicators
- **Education**: 3-4 indicators
- **Communication**: 2-3 indicators
- **Mortality**: 3-4 indicators
- **Deforestation**: 2-3 indicators

(Exact number varies by year and methodology)

### Temporal Comparisons

When comparing across years (2014, 2018, 2021, 2023):
- Methodology may have evolved between assessments
- New indicators may have been added
- Some municipalities may not have data in all years
- Use caution comparing very old (2014) with recent (2023) data

### Missing Data

- Some municipalities may lack data for specific indicators
- Remote or less accessible areas may have less complete data
- Use `na.rm = TRUE` in aggregations to handle missing values

### Geographic Coverage

- Covers 570+ municipalities in the Legal Amazon
- Includes all states with Amazon territory
- Some frontier/protected areas may lack complete data

***
