---
title: "Social Progress Index"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Social Progress Index}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The Amazônia Social Progress Index, an initiative from Imazon with support from 
the Social Progress Imperative is an index that measures the social and environmental
progress of its locations. Namely, the 772 municipalities in the Amazon region.
For more information, check the [initiative's website](http://www.ipsamazonia.org.br/conheca-o-ips).


## Usage

To download and compile data on the index for 2014 and 2018, just run the function
`load_IPS` using as its argument the directory in which the data should 
be stored. Choose with paramenter `language` whether yor output should be
in Portuguese or in English

```{r, eval = FALSE}
library(datazoom.amazonia)
my.data <- load_IPS(download_directory = getwd(), language = "pt")
```

## Credits

Data Zoom is developed by a team at Pontifícia Universidade Católica do Rio de Janeiro (PUC-Rio), Department of Economics. Our official website is at: http://www.econ.puc-rio.br/datazoom/index.html
