---
title: "MAPBIOMAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MAPBIOMAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

##The Dataset

The MAPBIOMAS project gathers data reporting the type of land covering each year by area, that is, for example, the area used for a temporary crop of soybeans. It also reports the transition between coverings during given years.

The data available has an yearly frequency and is available starting from the year 1989.

Using the function is easy enough:

```{r setup}
library(datazoom.amazonia)

data <- load_mapbiomas_cobertura(aggregation_level = 'municipality', path = NULL, code_state = "PA", code_mun = NULL, covering = 3, type = 'stacked', years = c(2000:2010))

data <- load_mapbiomas_transicao(path = NULL, code_state = "PA", covering_from = 3, covering_to = 19, type = 'normal', year_diff = 5)

```

Data can be retrieved from the website as above or loaded locally.
