---
title: "MAPBIOMAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MAPBIOMAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview

The MapBiomas project provides comprehensive satellite-based data on land cover and land use changes across Brazil. The project uses machine learning algorithms applied to Landsat satellite imagery to classify and track changes in biomes, land cover types, and economic activities over time. This dataset is essential for understanding deforestation, agricultural expansion, conservation outcomes, and landscape dynamics in the Amazon and other biomes.

MapBiomas represents one of the most detailed and scientifically rigorous sources of remote sensing data on Brazilian land cover, with coverage spanning from 1985 to present and annual updates.

## Data Coverage

The MapBiomas dataset includes:

- **Time Period**: Annual data from 1985 to 2023 (with updates continuing)
- **Geographic Coverage**: All Brazilian biomes (Amazon, Cerrado, Atlantic Forest, Caatinga, Pampa, Pantanal)
- **Spatial Resolution**: 30 meters (Landsat resolution)
- **Temporal Frequency**: Annual maps with seasonal detail
- **Update Schedule**: New data typically released annually with recent years added progressively
- **Aggregation Levels**: Municipalities, indigenous lands, states, biomes, and protected areas

## Available Datasets

### 1. Land Cover (`"mapbiomas_cover"`)

Annual maps of land cover types including:
- **Forest Types**: Native forest, forest formation, forest plantation
- **Non-Forest Vegetation**: Grassland/pasture, shrub vegetation, herbaceous
- **Agricultural Land**: Temporary crops (annual/seasonal), perennial crops (sugarcane, coffee, cocoa)
- **Urban Areas**: Urban/built-up land and infrastructure
- **Water Bodies**: Rivers, lakes, reservoirs, aquaculture
- **Non-vegetated**: Mining, bare soil, rock outcrops
- **Other Categories**: Cloud coverage, nodata

**Key Applications**: Deforestation monitoring, agricultural expansion, urbanization patterns, forest conservation

### 2. Land Cover Transitions (`"mapbiomas_transition"`)

Year-to-year changes in land cover including:
- **Deforestation**: Conversion from forest to other uses
- **Reforestation**: Conversion to forest types
- **Agricultural Transitions**: Changes between crop types or from other uses to agriculture
- **Degradation**: Forest to degraded/shrub vegetation
- **Regeneration**: Abandoned areas returning to vegetation

**Key Applications**: Deforestation tracking, regeneration assessment, agricultural dynamics, conservation effectiveness

### 3. Deforestation and Regeneration (`"mapbiomas_deforestation_regeneration"`)

Specific focus on forest cover changes:
- **Deforestation**: Permanent loss of forest cover
- **Forest Regeneration**: Secondary forest regrowth and natural recovery
- **Degradation Signals**: Forest areas showing stress indicators
- **Cumulative Deforestation**: Total forest loss since baseline

**Key Applications**: Amazon monitoring, conservation evaluation, regeneration potential, climate impact studies

### 4. Mining Activities (`"mapbiomas_mining"`)

Areas used for mining operations:
- **Active Mining**: Currently operational extraction sites
- **Abandoned Mining**: Previous mining areas, potentially available for rehabilitation
- **Mining Extent**: Total area impacted by mining activities
- **Mining Types**: Surface mining, artisanal mining, infrastructure

**Key Applications**: Environmental impact assessment, land degradation, restoration planning, environmental compliance

### 5. Irrigation (`"mapbiomas_irrigation"`)

**Note**: Temporarily unavailable - new collection coming soon

Previously included:
- **Irrigated Areas**: Extent of irrigation in agricultural systems
- **Irrigation Type**: Drip, center-pivot, flood irrigation
- **Crop Types Under Irrigation**: Which crops receive irrigation

### 6. Water Bodies (`"mapbiomas_water"`)

**Note**: Temporarily unavailable - new collection coming soon

Previously included:
- **Surface Water Extent**: Permanent and seasonal water bodies
- **Water Type**: Natural rivers/lakes vs. artificial reservoirs
- **Seasonal Variation**: Dry vs. wet season water extent

### 7. Wildfire Burn Scars (`"mapbiomas_fire"`)

Areas affected by wildfires:
- **Burn Scars**: Areas burned in recent fires
- **Fire Extent**: Total area impacted per year
- **Fire Frequency**: Repeated burning in same areas (indicating high fire risk)
- **Fire Season**: Temporal pattern of fires

**Key Applications**: Fire monitoring, ecosystem vulnerability, climate impacts, fire management

***

## Function Parameters

**Options:**

1. **dataset**: Seven possible choices (two temporarily unavailable)
   - `"mapbiomas_cover"`: Land cover types
   - `"mapbiomas_transition"`: Changes in land cover
   - `"mapbiomas_deforestation_regeneration"`: Forest cover changes
   - `"mapbiomas_mining"`: Mining areas
   - `"mapbiomas_irrigation"`: Irrigated areas (temporarily unavailable)
   - `"mapbiomas_water"`: Water bodies (temporarily unavailable)
   - `"mapbiomas_fire"`: Wildfire burn scars

2. **raw_data**: 
   - `TRUE`: Data in original format from MapBiomas
   - `FALSE`: Cleaned and standardized version (recommended for analysis)

3. **geo_level**: Varies by dataset
   - For `"mapbiomas_cover"`: `"municipality"` or `"indigenous_land"`
   - For `"mapbiomas_transition"`: `"municipality"` or `"biome"` (faster)
   - For `"mapbiomas_deforestation_regeneration"`: `"municipality"` only
   - For `"mapbiomas_mining"`: `"indigenous_land"` or `"municipality"`
   - For `"mapbiomas_irrigation"`: `"state"` or `"biome"`
   - For `"mapbiomas_water"`: `"municipality"`, `"state"`, or `"biome"`
   - For `"mapbiomas_fire"`: `"state"` only

4. **language**: 
   - `"pt"`: Portuguese language labels
   - `"eng"`: English language

***

## Measurement Units and Notes

- **All Areas**: Measured in hectares (ha)
- **1 hectare**: Approximately 2.47 acres
- **Validation**: MapBiomas uses field validation and visual interpretation to ensure accuracy
- **Updates**: Historical data may be revised with new collections and improved algorithms

***

***

## Practical Examples

### Example 1: Amazon Deforestation Trends

```{r eval=FALSE}
library(datazoom.amazonia)

# Download deforestation data for Amazon municipalities
deforestation <- load_mapbiomas(
  dataset = "mapbiomas_deforestation_regeneration",
  raw_data = FALSE,
  geo_level = "municipality",
  language = "eng"
)

# Identify municipalities with highest cumulative deforestation
top_deforested <- deforestation %>%
  group_by(municipality, state) %>%
  summarise(
    cumulative_deforestation = sum(deforestation_area, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(cumulative_deforestation)) %>%
  head(10)

print(top_deforested)
```

### Example 2: Land Cover Composition Analysis

```{r eval=FALSE}
# Download current land cover data at municipality level
land_cover <- load_mapbiomas(
  dataset = "mapbiomas_cover",
  raw_data = FALSE,
  geo_level = "municipality",
  language = "eng"
)

# Examine forest coverage by municipality (most recent year)
forest_cover <- land_cover %>%
  filter(year == max(year)) %>%
  filter(land_cover_type %in% c("Forest", "Forest Formation")) %>%
  group_by(municipality, state) %>%
  summarise(
    forest_area_ha = sum(area_hectares, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(forest_area_ha))

head(forest_cover, 15)
```

### Example 3: Agricultural Expansion Tracking

```{r eval=FALSE}
# Download land cover transitions data
transitions <- load_mapbiomas(
  dataset = "mapbiomas_transition",
  raw_data = FALSE,
  geo_level = "municipality",
  language = "eng"
)

# Track conversion from forest to agriculture
forest_to_ag <- transitions %>%
  filter(
    land_use_from %in% c("Forest", "Forest Formation"),
    land_use_to %in% c("Temporary Crops", "Sugarcane", "Pasture")
  ) %>%
  group_by(year, land_use_to) %>%
  summarise(
    conversion_area = sum(transition_area, na.rm = TRUE),
    .groups = 'drop'
  )

# Visualize agricultural expansion
plot(forest_to_ag$year, forest_to_ag$conversion_area,
     main = "Conversion from Forest to Agriculture",
     xlab = "Year", ylab = "Area (hectares)")
```

### Example 4: Biome-Level Analysis (Faster Download)

```{r eval=FALSE}
# Download transitions at biome level for faster processing
biome_transitions <- load_mapbiomas(
  dataset = "mapbiomas_transition",
  raw_data = FALSE,
  geo_level = "biome",  # Faster than municipality
  language = "pt"
)

# Compare deforestation rates across biomes
deforestation_by_biome <- biome_transitions %>%
  filter(land_use_from == "Forest") %>%
  group_by(biome, year) %>%
  summarise(
    annual_deforestation = sum(transition_area, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(biome, year)

print(deforestation_by_biome)
```

### Example 5: Mining Impact Assessment

```{r eval=FALSE}
# Download mining data by indigenous territory
mining_indigenous <- load_mapbiomas(
  dataset = "mapbiomas_mining",
  raw_data = FALSE,
  geo_level = "indigenous_land",
  language = "eng"
)

# Identify indigenous territories impacted by mining
impacted_territories <- mining_indigenous %>%
  filter(mining_area > 0) %>%
  arrange(desc(mining_area)) %>%
  head(15)

print(impacted_territories)
```

### Example 6: Fire Monitoring by State

```{r eval=FALSE}
# Download fire/burn scar data
fire_data <- load_mapbiomas(
  dataset = "mapbiomas_fire",
  raw_data = FALSE,
  geo_level = "state",
  language = "eng"
)

# Identify states with increasing fire activity
fire_trends <- fire_data %>%
  group_by(state) %>%
  summarise(
    total_burned_area = sum(burned_area, na.rm = TRUE),
    recent_fires = sum(burned_area[year >= 2020], na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(recent_fires))

head(fire_trends, 10)
```

### Example 7: Forest Regeneration Monitoring

```{r eval=FALSE}
# Download deforestation and regeneration data
forest_dynamics <- load_mapbiomas(
  dataset = "mapbiomas_deforestation_regeneration",
  raw_data = FALSE,
  geo_level = "municipality",
  language = "eng"
)

# Calculate net forest change and regeneration
forest_summary <- forest_dynamics %>%
  group_by(municipality, state) %>%
  summarise(
    total_deforestation = sum(deforestation_area, na.rm = TRUE),
    total_regeneration = sum(regeneration_area, na.rm = TRUE),
    net_change = total_regeneration - total_deforestation,
    .groups = 'drop'
  ) %>%
  filter(net_change > 0) %>%  # Municipalities with net forest gain
  arrange(desc(net_change))

# Identify regeneration success stories
print(head(forest_summary, 10))
```

### Example 8: Temporal Trend Analysis

```{r eval=FALSE}
# Download multi-year land cover data
long_series <- load_mapbiomas(
  dataset = "mapbiomas_cover",
  raw_data = FALSE,
  geo_level = "municipality",
  language = "eng"
)

# Track one municipality over time
specific_municipality <- long_series %>%
  filter(municipality == "Altamira" & state == "PA") %>%
  group_by(year, land_cover_type) %>%
  summarise(area = sum(area_hectares, na.rm = TRUE), .groups = 'drop')

# Visualize landscape change
library(ggplot2)
ggplot(specific_municipality, aes(x = year, y = area, fill = land_cover_type)) +
  geom_area() +
  theme_minimal() +
  labs(title = "Land Cover Changes in Altamira (PA)",
       x = "Year", y = "Area (hectares)")
```

***

## Use Cases

- **Deforestation Monitoring**: Track Amazon forest loss and environmental policies effectiveness
- **Agricultural Planning**: Monitor crop expansion and land use intensity
- **Environmental Compliance**: Verify land use against environmental regulations
- **Conservation Assessment**: Evaluate protected area effectiveness and forest regeneration
- **Climate Impact**: Estimate carbon implications of land use changes
- **Restoration Planning**: Identify areas suitable for reforestation or regeneration
- **Mining Impact**: Assess environmental damage from mining operations
- **Fire Management**: Monitor fire patterns and ecosystem vulnerability

***
