---
title: "MAPBIOMAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MAPBIOMAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The MAPBIOMAS project gathers data reporting the type of land covering each year by area, that is, for example, the area used for a temporary crop of soybeans. It also reports the transition between coverings during given years.

The data available has an yearly frequency and is available starting from the year 1989.

For all datasets the unit of measurement is hectares(ha). 

***
**Options:**

  1. **dataset**:
     * `"mapbiomas_cover"`: types of land cover.
     * `"mapbiomas_transition"`: changes in land cover.
     * `"mapbiomas_deforestation_regeneration"`: deforestation and forest regeneration.
     * `"mapbiomas_irrigation"`: irrigated areas.
     * `"mapbiomas_mining"`: areas used for mining.
     * `"mapbiomas_water"`: areas of water surface.
     * `"mapbiomas_fire"`: areas of wildfire burn scars.


  2. **raw_data**: there are two options:
     * `TRUE`: if you want the data as it is originally.
     * `FALSE`: if you want the treated version of the data. 


  3. **geo_level**:
     * For datasets `"mapbiomas_cover"`or `"mapbiomas_transition"` can be `"municipality"` or `"state"` (faster download).
     * For dataset `"mapbiomas_deforestation_regeneration"` can be only `"municipality"`.
     * For dataset `"mapbiomas_mining"`, can be `"indigenous_land"` or `"municipality"`.
     * For dataset `"mapbiomas_irrigation"`, can be `"state"` or `"biome"`.
     * For dataset `"mapbiomas_water"`, can be `"municipality"`, `"state"` or `"biome"`.
     * For dataset `"mapbiomas_fire"`, can be only `"state"`.
     * Does not apply to other datasets.


  4. **language**: you can choose between Portuguese `("pt")` and English `("eng")`.
  
***

**Examples:** 
 
```{r eval=FALSE}
# download treated Mapbiomas Cover data in english
data <- load_mapbiomas(dataset = "mapbiomas_cover",
                       raw_data = FALSE,
                       geo_level = "municipality",
                       language = "eng")

# download treated Mapbiomas Transition data in portuguese
data <- load_mapbiomas(dataset = "mapbiomas_transition", raw_data = FALSE,
                       geo_level = "state", language = "pt")

# download treated data on mining on indigenous lands
data <- load_mapbiomas("mapbiomas_mining",
                       raw_data = FALSE,
                       geo_level = "indigenous_land")
```


