---
title: "MAPBIOMAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MAPBIOMAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The MAPBIOMAS project gathers data reporting the type of land covering each year by area, that is, for example, the area used for a temporary crop of soybeans. It also reports the transition between coverings during given years.

The data available has an yearly frequency and is available starting from the year 1989.
 
***
**Options:**

  1. **dataset**:
      * `"mapbiomas_cover"`: types of land cover
      * `"mapbiomas_transition"`: changes in land cover
      * `"mapbiomas_deforestation_regeneration"`: deforestation and forest regeneration
      * `"mapbiomas_irrigation"`: irrigated areas
      * `"mapbiomas_grazing_quality"`: grazing quality
      * `"mapbiomas_mining"`: areas used for mining

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. 
      
  3. **geo_level**:
      * For datasets `"mapbiomas_cover"` and `"mapbiomas_transition"`, can be `"municipality"` or `"state"` (faster download).
      * For dataset `"mapbiomas_mining"`, can be `"biome"` or `"indigenous_land"`.
      * Does not apply to other datasets.

  4. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
  5. **cover_level**: Aggregates the data to some level of land coverage. Only applies to datasets `"mapbiomas_cover"` and `"mapbiomas_grazing_quality"`:
      * `cover_level = "none"`: no aggregation
      * `cover_level = 0`: least aggregated, with categories of Anthropic and Natural
      * `cover_level = 1`: categories such as Forest, Non Forest Natural Formation, Farming, Non Vegetated Area, Water, Non Observed
      * `cover_level = 2`: categories such as Agriculture, Aquaculture, Beach and Dune, Forest Plantation,  Pasture, River, Lake and Ocean                            
      * `cover_level = 3`: categories such as Aquaculture, Beach and Dune, Forest Formation, Forest Plantation 
      * `cover_level = 4`: categories such as Aquaculture, Beach and Dune, Forest Formation, Forest Plantation 
  
***

**Examples:** 
 
```{r eval=FALSE}
# download treated Mapbiomas Cover data in english at the highest aggregation level
data <- load_mapbiomas(dataset = "mapbiomas_cover",
                       raw_data = FALSE,
                       geo_level = "municipality",
                       language = "eng",
                       cover_level = 0)

# download treated Mapbiomas Transition data in portuguese
data <- load_mapbiomas(dataset = "mapbiomas_transition", raw_data = FALSE,
                       geo_level = "state", language = "pt")

# download treated data on mining on indigenous lands
data <- load_mapbiomas("mapbiomas_mining",
                       raw_data = FALSE,
                       geo_level = "indigenous_land")
```


