---
title: "legal_amazon"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{legal_amazon}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `municipalities` dataset is a foundational reference dataset included in the datazoom.amazonia package. It contains key information about all Brazilian municipalities and identifies which are located in (or overlapping with) the Legal Amazon region.

This dataset includes:

- **Municipality identification**: Official names, codes (IBGE), and geographic identifiers
- **Legal Amazon status**: Identifies whether each municipality is in the Legal Amazon
- **Geographic information**: State, region, and spatial boundaries
- **Supporting attributes**: Additional useful variables for geographic matching and analysis
- **Historical coverage**: Updated to reflect current municipal boundaries (2019 onwards)

The municipalities dataset is essential infrastructure for this package—most package functions that return geographic data match results to this municipalities reference to enable Legal Amazon filtering and consistent geographic identification.

### Data Source

The municipalities dataset is compiled from:
- **IBGE**: Brazilian Institute of Geography and Statistics official municipal data
- **Legal Amazon definition**: Based on official Brazilian government definition
- **Spatial boundaries**: IBGE 2019 municipal boundary shapefile
- **Maintained by**: datazoom.amazonia package developers

***

## Dataset Structure

The municipalities dataset contains the following information:

### Key Variables

- **code**: IBGE municipal code (unique identifier)
- **name**: Official municipality name
- **state**: Two-letter state abbreviation (e.g., "SP", "AM")
- **state_code**: IBGE state code
- **legal_amazon**: TRUE/FALSE indicating Legal Amazon membership
- **region**: Geographic region (North, Northeast, Center-West, Southeast, South)
- **geometry**: Spatial polygon (when using full SF object)

### Data Types

- **IBGE codes**: Numeric identifiers used throughout Brazilian statistics
- **State abbreviations**: Standard two-letter codes for all 27 Brazilian states/federal district
- **Region names**: Official Brazilian geographic regions
- **Geographic boundaries**: Available as simple features (SF) objects for spatial analysis

***

## Accessing the Dataset

### In R Code

```{r eval = FALSE}
# Load Brazilian municipalities dataset
data <- datazoom.amazonia::municipalities

# Or after loading the package
library(datazoom.amazonia)
data <- municipalities

# View structure
str(municipalities)
head(municipalities)

# Filter for Legal Amazon municipalities only
amazon_municipalities <- municipalities %>%
  filter(legal_amazon == TRUE)
```
```
### What It Contains

The dataset includes all 5,570+ Brazilian municipalities with:
- Official IBGE identification codes
- State and region classification
- Legal Amazon flag for filtering
- Spatial geometries for geographic analysis

---
## Common Use Cases

### Use Case 1: Filter Data to Legal Amazon

Many analyses focus specifically on the Legal Amazon region. Use the municipalities dataset to identify relevant municipalities:

```{r eval=FALSE}
library(dplyr)
# Load any dataset with municipality information
data <- load_prodes(
  dataset = "deforestation",
  raw_data = FALSE,
  geo_level = "municipality",
  language = "eng"
)

# Filter to Legal Amazon using municipalities reference
amazon_data <- data %>%
  inner_join(
    municipalities %>% 
      filter(legal_amazon == TRUE) %>%
      select(code, name),
    by = c("municipality_code" = "code")
  )
```

### Use Case 2: Partial Amazon Municipalities

 **Important**: Some municipalities are only partially within the Legal Amazon.

For statistics reported at municipality level in this package:
- **Partial Amazon municipalities**: Data is reported for only the Amazon-included portion
- **Identification**: The municipalities dataset identifies these cases
- **Interpretation**: When a municipality is partially in Amazon, reported statistics reflect the Amazon portion only

```{r eval=FALSE}
# Identify fully vs. partially included municipalities
full_amazon <- municipalities %>%
  filter(legal_amazon == TRUE & fully_included == TRUE)

partial_amazon <- municipalities %>%
  filter(legal_amazon == TRUE & fully_included == FALSE)

print(paste("Fully in Amazon:", nrow(full_amazon)))
print(paste("Partially in Amazon:", nrow(partial_amazon)))
```

***

## Key Characteristics

### Municipal Coverage

- **Total municipalities**: 5,570+ (exact number updated periodically)
- **Geographic coverage**: All Brazilian territory
- **Boundaries**: Based on IBGE 2019 definitions (latest update)
- **Changes**: Municipal boundaries periodically updated (most recent: 2021)

### Legal Amazon Definition

The Legal Amazon includes:
- **States**: All or parts of Acre, Amapá, Amazonas, Distrito Federal, Goiás, Maranhão, Mato Grosso, Mato Grosso do Sul, Pará, Rondônia, Roraima, Tocantins
- **Municipalities**: 570+ municipalities fully or partially in Legal Amazon
- **Definition**: Based on official Brazilian legislation (Law 8,001/1990)

### Partial Inclusion

Important note on municipality-level data:
- **Partial municipalities**: Some municipalities extend beyond Legal Amazon boundaries
- **Data reporting**: When data is reported at municipality level for partial municipalities, it reflects **only the Legal Amazon portion**
- **Identification**: This dataset identifies which municipalities are partial

```{r eval=FALSE}
# Check for partial municipalities
partial_check <- municipalities %>%
  filter(legal_amazon == TRUE) %>%
  filter(!is.na(amazon_percentage)) %>%
  filter(amazon_percentage < 100)

if (nrow(partial_check) > 0) {
  print("Municipalities partially in Legal Amazon:")
  print(partial_check)
}
```

***
***

## Important Notes

### Boundary Changes

Brazilian municipalities occasionally undergo changes:
- **New municipalities**: Created from existing ones (last major change 2021)
- **Boundary adjustments**: IBGE periodically refines municipal boundaries
- **Historical data**: When comparing very old data with recent data, be aware municipalities may have been reorganized

### Code Consistency

Always use IBGE municipality codes (not names) when:
- Merging multiple datasets
- Doing time-series analysis
- Comparing across sources
- Municipality names may change or be ambiguous; codes are unique and stable

### Spatial Data

The municipalities dataset includes spatial boundaries (when loaded as SF object):
- **Format**: Simple features (SF) polygons
- **CRS**: WGS84 (EPSG:4326)
- **Use**: Spatial operations, mapping, spatial joins with other geographic data

***



## Accessing Spatial Data

### For Mapping and Spatial Analysis

```{r eval=FALSE}
library(sf)
library(ggplot2)

# Load municipalities with geometry
municipalities_sf <- municipalities %>%
  st_as_sf()  # If not already SF format

# Map Legal Amazon
amazon_map <- municipalities_sf %>%
  filter(legal_amazon == TRUE)

ggplot(amazon_map) +
  geom_sf(fill = "lightgreen", color = "darkgreen") +
  labs(title = "Legal Amazon Municipalities") +
  theme_minimal()

# Spatial operations example: count municipalities by state
munic_by_state <- municipalities_sf %>%
  group_by(state) %>%
  summarize(
    num_municipalities = n(),
    total_area_km2 = sum(st_area(.), na.rm = TRUE) / 1e6,
    .groups = 'drop'
  )
```

***
***

## Troubleshooting

### Matching Issues

**Problem**: Municipality names don't match between datasets
**Solution**: Use IBGE municipality codes instead of names for joining data

**Problem**: Some municipalities missing after filtering
**Solution**: Check for name spelling variations; use code-based matching

### Geographic Analysis

**Problem**: Spatial operations are slow
**Solution**: Simplify geometries (`st_simplify()`) or work with state/region level first

**Problem**: Mapping appears incorrect
**Solution**: Verify CRS (should be WGS84); check for invalid geometries (`st_is_valid()`)

### Data Integration

**Problem**: Aggregating municipality data to regions
**Solution**: Use `left_join()` with municipalities dataset to add region information

**Problem**: Partial municipalities causing data discrepancies
**Solution**: Account for partial Amazon municipalities; some statistics are reported for Amazon portion only

***

## Related Resources

- **IBGE**: https://www.ibge.gov.br/ (Brazilian statistics authority)
- **Legal Amazon**: Official definition https://www.gov.br/
- **Spatial data**: IBGE municipal boundaries available from various sources
- **Within this package**: All geographic functions integrate this municipalities reference

```
```


