---
title: "PEVS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PEVS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

Silviculture and Forestry Extraction Production (PEVS, in Portuguese) is a nationwide annual survey conducted by IBGE (Brazilian Institute of Geography and Statistics) which provides information on the quantity and value of production resulting from the processes of exploitation of native plant resources and planted forest massifs. 

The data also has multiple aggregation levels, such as nationwide, by region, mesoregion and microregion, as well as state and municipality. 

The data available has a yearly frequency and is available from 1986 to the present, with the exception of the data on total area for production, which are only available from 2013 onwards. More information can be found on this link (only in Portuguese): https://www.ibge.gov.br/estatisticas/economicas/agricultura-e-pecuaria/9105-producao-da-extracao-vegetal-e-da-silvicultura.html?=&t=o-que-e

## Usage

<!-- The package contains of three function: \code{load_pevs_vegextr}, \code{load_pevs_silvprod} and \code{load_pevs_areasilv} which extract data from the API of the IBGE database website.  -->

<!-- The first function provides data from products of extractivism, which means resources extracted from the forest that were not originally planted by humans, such as dyes, waxes, fibers, medicinal and aromatic herbs, and much more. Products are grouped in these categories, and for each product extracted there is information on the quantity produced and the monetary value of this yield (Brazil currency has changed over the years so, depending on the period, the currency name may change). -->

<!-- The second function provides data from products of silviculture, which is the practice of cultivating forest crops to meet the needs of wood production for uses such as firewood, charcoal or paper and celulose production. Similarly to the first function, for each product there is data on the quantity produced and the monetary value of this yield. Products are grouped by usage categories, with the main products being eucalyptus and pine. -->

<!-- The third function, however, pertains to the total existing area, in hectares, used for silviculture as of December 31st of each year. This survey is more recent, and the products available are eucalyptus, pine and a third category of "other species".  -->

<!-- All of the functions allow the user to input the years he would like the data to encompass, the geographical aggregation level desired, as well as choosing the language of the data (currently only Portuguese and English are supported). These functions also offer the option to output the data in the long format, in which separate tables are created for each variable and each product becomes a column (whereas previously both were lines of one single larger data frame). -->

<!-- The output is a data frame containing data for the territorial level chosen, the name of such geographic level, the year of the data, the variable in question, the unit of measurement and the value. Each column has a corresponding column containing the numerical code of the specific value, with the exception of the Value column, which is already numeric. If the long format is selected, the output becomes a list of data frames, one for each one of the variables. -->

Using the function is easy enough:

```{r eval = FALSE}
library(datazoom.amazonia)

# download state raw data from 2012 for silviculture
pevs_silvi <- load_pevs(dataset = 'pevs_silviculture',
                        raw_data = TRUE,
                        geo_level = 'state',
                        time_period = 2012)
```
