---
title: "PIBMUNIC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PIBMUNIC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

PIBMUNIC (Produto Interno Bruto por Município - Gross Domestic Product by Municipality) is Brazil's official municipal-level GDP data produced by [IBGE](https://www.ibge.gov.br/). This dataset provides:

- **Gross Domestic Product (GDP)**: Total economic output at current market prices
- **Value Added**: Gross value added by economic activity
- **Taxes and Subsidies**: Net taxes on products
- **Sectoral breakdown**: GDP disaggregated by economic sectors (agriculture, industry, services)
- **Municipal coverage**: All Brazilian municipalities
- **Long time series**: Historical data spanning multiple decades
- **Multi-level aggregation**: Available at country, state, and municipality levels

PIBMUNIC is essential for understanding Brazil's regional economic structure, identifying economic disparities, analyzing sectoral specialization, and assessing economic development across municipalities.

### Data Source and Methodology

PIBMUNIC data is compiled by IBGE using:
- Data from CEMPRE (firm registry) for employment and output
- Production and consumption surveys
- Tax and financial records
- Trade and services data
- National accounts framework aligned with international standards

For more information, visit [IBGE National Accounts](https://www.ibge.gov.br/en/statistics/economic/national-accounts/).

***

## Available Dataset

### **pibmunic**

Complete municipal GDP statistics with sectoral detail.

- **Coverage**: All 5,570+ Brazilian municipalities
- **Variables**: GDP at current prices, value added by sector, taxes, subsidies
- **Sectors included**: Agriculture, industry, services, public administration
- **Measurement**: Brazilian Real (R$) at current prices
- **Time period**: Varies by year; typically 2002 onwards
- **Use cases**:
  - Identify richest and poorest municipalities
  - Analyze regional economic disparities
  - Assess sectoral specialization (agriculture vs. industry vs. services)
  - Economic growth analysis by municipality
  - Development planning and policy evaluation
  - Correlate with social/environmental indicators

***

## Function Parameters

### 1. **dataset**

Only one dataset is available:

```r
dataset = "pibmunic"  # Municipal GDP data
```

### 2. **raw_data**

Controls whether to download original or processed data.

- `TRUE`: Returns raw IBGE format
- `FALSE`: Returns treated data with English variable names and standardized formatting

```r
raw_data = FALSE  # logical
```

### 3. **geo_level**

Specifies geographic aggregation level.

- `"country"`: National aggregate
- `"state"`: State-level aggregation
- `"municipality"`: Most detailed level with all municipalities

```r
geo_level = "municipality"  # character string
```

### 4. **time_period**

Specifies which year(s) to download.

```r
time_period = 2020              # single year
time_period = c(2015, 2020)     # specific years
time_period = 2015:2020         # range of years
```

### 5. **language**

Output language for variable names.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

***

## Examples

### Example 1: Municipal GDP for a single year

```{r eval=FALSE}
# download treated municipal GDP data for 2020
pib_munic <- load_pibmunic(
  dataset = "pibmunic",
  raw_data = FALSE,
  geo_level = "municipality",
  time_period = 2020,
  language = "eng"
)
```

### Example 2: State-level GDP over time

```{r eval=FALSE}
# download treated state-level GDP data for 2015 to 2020
pib_state <- load_pibmunic(
  dataset = "pibmunic",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2015:2020,
  language = "eng"
)
```

### Example 3: Country-level GDP in Portuguese

```{r eval=FALSE}
# download treated country-level GDP data for 2010 to 2020 in Portuguese
pib_br <- load_pibmunic(
  dataset = "pibmunic",
  raw_data = FALSE,
  geo_level = "country",
  time_period = 2010:2020,
  language = "pt"
)
```

## Data Notes

### Variables Structure

Typical variables include:
- **gdp**: Gross Domestic Product at current prices (R$)
- **value_added_agriculture**: Farming and forestry sector
- **value_added_industry**: Manufacturing and construction
- **value_added_services**: Commerce, finance, transport, etc.
- **value_added_public_admin**: Government services
- **net_taxes**: Taxes minus subsidies
- Additional detail depends on specific IBGE release

### Current Prices

- All values are in "current prices" (nominal values, not deflated)
- For real (inflation-adjusted) comparisons, you need to apply price deflators
- Comparisons across years should account for inflation

### Data Coverage

- **All municipalities**: Includes all 5,570+ Brazilian municipalities
- **Time lag**: Data typically released 2+ years after reference year
- **Revisions**: IBGE periodically revises historical data

### Important Limitations

1. **Current prices**: Values not adjusted for inflation
2. **Time lag**: Recent years may not be available
3. **Confidentiality**: Some small municipalities may have aggregated data
4. **Methodological changes**: IBGE occasionally updates national accounts methodology
5. **Municipal boundaries**: Changed in 2021; affects historical comparisons

***

