---
title: "POPULATION"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{POPULATION}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Loads IBGE information on estimated population (2001-2006, 2008-2009, 2011-2021) or population (2007 and 2010) data. Data is available at country, state and municipality level and from 2001 to 2021.

***
**Options:**

  1. **dataset**: there are 6 possible options:
  
      * `"resident_estimated"`: Resident population in the Demographic Census.
      * `"resident_race_census"`: Resident population, by color or race, in the Demographic Census.
      * `"resident_race_age"`: Resident population, by color or race, according to sex and age.
      * `"aging_race"`: Aging index, median age, and sex ratio, by color or race.
      * `"indigenous_age_sex"`: Indigenous population, by age groups and sex.
      * `"aging_indigenous"`: Aging index, median age, and sex ratio of the indigenous population, according to the declaration criterion.
      

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. 
      
  3. **geo_level**: `"country"`, `"state"`, or `"municipality"`    
  
  4. **time_period**: picks the years for which the data will be downloaded

  5. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
***

**Examples:**

```{r, eval = FALSE}
# download treated population data at the state level for 2010 to 2012
data <- load_population(
  dataset = "resident_estimated",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2010:2012
)
```

