---
title: "POPULATION"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{POPULATION}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview

Population dataset provides Brazil's official population statistics from [IBGE](https://www.ibge.gov.br/) (Brazilian Institute of Geography and Statistics). This dataset includes:

- **Population estimates**: Official projections and estimates for non-census years
- **Census data**: Actual population counts from census years (2007, 2010)
- **Time coverage**: Data from 2001 to 2021 (with some gaps)
- **Geographic detail**: Available at country, state, and municipality levels
- **Regular updates**: Annual estimates published by IBGE
- **Demographic foundation**: Base data for per capita calculations and demographic analysis

Population data is fundamental to many analyses, providing denominators for per capita metrics, understanding urbanization patterns, tracking demographic trends, and supporting policy and development planning.

### Data Source and Methodology

Population data comes from:
- **Census years (2007, 2010)**: Actual population enumeration
- **Estimation years (2001-2006, 2008-2009, 2011-2021)**: IBGE official population projections based on census data and vital statistics
- **Methodology**: Cohort-component methodology considering births, deaths, migration
- **Updates**: Revised annually as new vital statistics data becomes available

For more information, visit [IBGE Population Estimates](https://www.ibge.gov.br/en/statistics/social/population/).

---

## Available Dataset

### **population**

Official Brazilian population statistics and estimates.

- **Coverage**: All Brazilian municipalities, states, and national level
- **Time period**: 2001 to 2021
- **Data types**: Census counts (2007, 2010) and official estimates (other years)
- **Geographic levels**: Country, State, Municipality
- **Use cases**:
  - Calculate per capita indicators for all metrics
  - Track demographic changes and urbanization
  - Assess population concentration
  - Population-weighted aggregations
  - Demographic planning and projections
  - Support for epidemiological and social statistics

---

## Important Data Characteristics

### Census Years vs. Estimate Years

| Year Type | Years | Nature |
|-----------|-------|--------|
| Census | 2007, 2010 | Actual population count |
| Estimates | 2001-2006, 2008-2009, 2011-2021 | Official IBGE projections |

Census years provide actual counts; other years are official projections.

### Population Definition

- **Resident population**: People living in a given area as of the reference date
- **Includes**: Brazilian citizens and legal residents
- **Reference date**: July 1 for most estimate years, September 1 for 2010 census

---

## Function Parameters

### 1. **dataset**

Only one dataset is available:

```r
dataset = "population"  # Population statistics
```

### 2. **raw_data**

Controls whether to download original or processed data.

- `TRUE`: Returns raw IBGE data format
- `FALSE`: Returns treated data with English variable names and standardized format

```r
raw_data = FALSE  # logical
```

### 3. **geo_level**

Specifies geographic aggregation level.

- `"country"`: National total
- `"state"`: State-level population (27 units)
- `"municipality"`: All 5,570+ municipalities

```r
geo_level = "municipality"  # character string
```

### 4. **time_period**

Specifies which year(s) to download.

**Available years**: 2001-2006, 2007 (census), 2008-2009, 2010 (census), 2011-2021

```r
time_period = 2020              # single year
time_period = c(2010, 2020)     # specific years (includes 2010 census)
time_period = 2010:2020         # range of years
```

### 5. **language**

Output language for variable names.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

---

## Examples

### Example 1: Population by state for a single year

```{r eval=FALSE}
# download treated population data at the state level for 2021
pop_2021 <- load_population(
  dataset = "population",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2021,
  language = "eng"
)
```

### Example 2: Population by state over time

```{r eval=FALSE}
# download treated population data at the state level for 2010 to 2021
pop_series <- load_population(
  dataset = "population",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2010:2021,
  language = "eng"
)
```

### Example 3: Population by municipality

```{r eval=FALSE}
# download treated population data at the municipality level for 2020
pop_munic <- load_population(
  dataset = "population",
  raw_data = FALSE,
  geo_level = "municipality",
  time_period = 2020,
  language = "eng"
)
```

## Data Notes

### Data Structure

Each record contains:
- Geographic identifier (state or municipality name/code)
- Year
- Population count

Simple and straightforward structure, useful as base for other calculations.

### Census vs. Estimate Quality

- **Census years (2007, 2010)**: High quality, actual counts
- **Estimate years**: Official IBGE projections based on demographic models
- **Estimates are less certain**: Especially for small geographic areas

### Municipal Variations

- **Not all municipalities equally reliable**: Very small municipalities may have estimation uncertainty
- **Boundary changes**: Municipalities occasionally merged or divided; affects time comparisons
- **Code consistency**: Use IBGE municipal codes (rather than names) for accurate merging with other datasets

### Important Limitations
1. **Estimate years**: Non-census years are projections, not actual counts
2. **Lag in release**: May take time for latest estimates to be published
3. **Revisions**: Estimates revised annually as better vital statistics data becomes available
4. **Boundary changes**: Municipal boundaries changed; affects historical comparisons
5. **Definition**: Includes registered residents; may not match actual lived population

---

