---
title: "PPM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PPM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Data on livestock inventories (e.g:cattle, pigs and hogs) in Brazilian Municipalities, as well as amount and value of animal products (e.g:output of milk, hen eggs, quail eggs, honey).

The periodicity of the survey is annual. The geographic coverage is national, with results released for Brazil, Major Regions, Federation Units, Mesoregions, Microregions and Municipalities.

The data available has a yearly frequency and is available from 1974 to the present. More information can be found in [this link](https://www.ibge.gov.br/en/statistics/economic/agriculture-forestry-and-fishing/17353-municipal-livestock-production.html?=&t=o-que-e).

***
**Options:**

  1. **dataset**: there are five possible choices:
  
      * `"ppm_livestock_inventory"`: livestock herds, with the number of animals by species.
      
      * `"ppm_sheep_farming"`: total sheep reared and specific shearing data.
      
      * `"ppm_animal_origin_production"`: animal-origin products like milk, eggs, and honey, as well as wool and other derivatives.
      
      * `"ppm_cow_farming"`: dairy cow farming, with the number of milked cows, their geographic distribution, and productivity rates.
      
      * `"ppm_aquaculture"`: aquaculture activities, including fish farming, shrimp farming, and mollusk farming.

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. 
      
  3. **geo_level**: `"country"`, `"region"`, `"state"`, or `"municipality"`    
  
  4. **time_period**: picks the years for which the data will be downloaded

  5. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
***

**Examples:**

```{r eval=FALSE}
# Download treated data (raw_data = FALSE) about aquaculture (dataset = "ppm_aquaculture")
# from 2013 to 2015 (time_period = 2013:2015) in english
# with the level of aggregation being the country (geo_level = "country").
data <- load_ppm(
  dataset = "ppm_aquaculture",
  raw_data = FALSE,
  geo_level = "country",
  time_period = 2013:2015
)

# Download raw data about sheep farming by state from 1980 to 1995 in portuguese (language = "pt")
data <- load_ppm(
  dataset = "ppm_sheep_farming",
  raw_data = TRUE,
  geo_level = "state",
  time_period = 1980:1995,
  language = "pt"
)
```
