---
title: "PPM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PPM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Data on livestock inventories (e.g:cattle, pigs and hogs) in Brazilian Municipalities, as well as amount and value of animal products (e.g:output of milk, hen eggs, quail eggs, honey).

The periodicity of the survey is annual. The geographic coverage is national, with results released for Brazil, Major Regions, Federation Units, Mesoregions, Microregions and Municipalities.

The data available has a yearly frequency and is available from 1974 to the present. More information can be found in [this link](https://www.ibge.gov.br/en/statistics/economic/agriculture-forestry-and-fishing/17353-municipal-livestock-production.html?=&t=o-que-e).

```
There are five parameters in this function:
  
  1. dataset: There are five possible choices. 
  # 'ppm_livestock_inventory'
  # 'ppm_sheep_farming'
  # 'ppm_animal_orig_production' 
  # 'ppm_cow_farming'
  # 'ppm_aquaculture' 
  2. raw_data: there are two options:
  # TRUE: if you want the data as it is in the IBGE's site.
  # FALSE: if you want the treated (more organized) version of the data. 
  3. geo_level: there are four options:
  # "country"
  # "region"
  # "state"
  # "municipality"
    
  4. time_period:
  
  # For ppm_livestock_inventory, ppm_sheep_farming, ppm_animal_orig_production and ppm_cow_farming: data is avaiable from 1974 to 2019.
  # For ppm_aquaculture: data is avaiable from 2013 to 2019 
  
  5. language: you can choose between portuguese and english
```  

```{r eval=FALSE}
# Download treated data (raw_data = FALSE) about aquaculture (dataset = 'ppm_aquaculture') from 2013 to 2015 (time_period = 2013:2015) in english (language = "eng") with the level of aggregation being the country (geo_level = "country"). 
data <- load_ppm(dataset = 'ppm_aquaculture', 
                 raw_data = FALSE, 
                 geo_level = "country", 
                 time_period = 2013:2015, 
                 language = "eng")
# Download raw data (raw_data = TRUE) about sheep farming (dataset = 'ppm_sheep_farming') by state (geo_level = "state") from 1980 to 1995 (time_period = 1980:1995) in portuguese (language = "pt")
data <- load_ppm(dataset = 'ppm_sheep_farming', 
                 raw_data = TRUE, 
                 geo_level = "state", 
                 time_period = 1980:1995, 
                 language = "pt")
```
