---
title: "PPM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PPM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

PPM (Pesquisa da Pecuária Municipal - Municipal Livestock Survey) is Brazil's comprehensive annual survey of livestock activities conducted by [IBGE](https://www.ibge.gov.br/). This dataset provides:

- **Livestock inventories**: Number of animals by species (cattle, pigs, poultry, sheep, horses, etc.)
- **Animal products**: Production volumes and values of milk, eggs, honey, wool, and other animal-origin products
- **Dairy farming**: Specialized data on milked cows, geographic distribution, productivity
- **Aquaculture**: Fish farming, shrimp farming, and mollusk farming activities
- **Sheep specialization**: Detailed shearing and wool production data
- **Multi-level geographic detail**: Country, region, state, and municipality levels
- **Long historical series**: Available from 1974 onwards
- **Economic value**: Both production quantities and market values

PPM is the primary data source for understanding Brazil's livestock sector, which is economically significant and globally important for beef, poultry, and dairy exports.

### Data Source and Methodology

PPM data is compiled from:
- Direct surveys of livestock producers
- Agricultural censuses and administrative records
- Municipal agriculture secretariats
- Processed and validated by IBGE
- Annual release with data for reference year

For more information, visit [IBGE Livestock Statistics](https://www.ibge.gov.br/en/statistics/economic/agriculture-forestry/).

***

## Available Datasets

### **1. ppm_livestock_inventory**

Total livestock herds disaggregated by animal species.

- **Coverage**: All livestock species across all Brazilian municipalities
- **Time period**: 1974 onwards
- **Geographic levels**: Country, Region, State, Municipality
- **Animal species**: Cattle, pigs, chickens, sheep, horses, goats, buffalo, others
- **Variables**: Number of animals by species, number of establishments
- **Use cases**:
  - Identify regional livestock specialization
  - Track herd size trends
  - Analyze geographic concentration of livestock
  - Understand animal agriculture structure

### **2. ppm_sheep_farming**

Specialized data on sheep production and wool/fleece harvest.

- **Coverage**: Sheep farming across Brazil
- **Time period**: 1974 onwards
- **Geographic levels**: Country, Region, State, Municipality
- **Variables**: Total sheep, sheared sheep, fleece weight, wool production
- **Use cases**:
  - Analyze wool production and sheep farming specialization
  - Track shearing practices and yields
  - Regional wool industry assessment

### **3. ppm_animal_origin_production**

Production of animal-based products (milk, eggs, honey, wool, etc.).

- **Coverage**: All animal product production activities
- **Time period**: 1974 onwards
- **Geographic levels**: Country, Region, State, Municipality
- **Products included**: Cow milk, goat milk, chicken eggs, quail eggs, honey, wool, hides, wax
- **Variables**: Quantity produced and value of production
- **Use cases**:
  - Track dairy and egg production
  - Analyze honey and other bee products
  - Economic analysis of animal product sectors

### **4. ppm_cow_farming**

Detailed dairy cow farming data with milking and productivity metrics.

- **Coverage**: Dairy cow operations
- **Time period**: 1974 onwards
- **Geographic levels**: Country, Region, State, Municipality
- **Variables**: Milked cows, milk production volume, productivity (liters per cow)
- **Use cases**:
  - Dairy sector analysis
  - Productivity assessment
  - Geographic specialization in dairy
  - Production trend analysis

### **5. ppm_aquaculture**

Aquaculture activities including fish, shrimp, and mollusk farming.

- **Coverage**: All aquaculture operations
- **Time period**: 1974 onwards (though aquaculture data more recent)
- **Geographic levels**: Country, Region, State, Municipality
- **Activities**: Fish farming, shrimp farming, mollusk/oyster farming, other aquaculture
- **Variables**: Quantity and value of aquaculture production by type
- **Use cases**:
  - Aquaculture sector analysis
  - Regional aquaculture potential
  - Fish and seafood production trends
  - Emerging aquaculture development

***

## Function Parameters

### 1. **dataset**

Selects which livestock/animal production dataset to download.

```r
dataset = "ppm_livestock_inventory"     # Animal populations by species
dataset = "ppm_sheep_farming"           # Sheep and wool production
dataset = "ppm_animal_origin_production"  # Milk, eggs, honey, wool
dataset = "ppm_cow_farming"             # Dairy cow productivity
dataset = "ppm_aquaculture"             # Fish and aquaculture production
```

### 2. **raw_data**

Controls whether to download original or processed data.

- `TRUE`: Returns raw IBGE format
- `FALSE`: Returns treated data with English variable names and standardized units

```r
raw_data = FALSE  # logical
```

### 3. **geo_level**

Specifies geographic aggregation level.

- `"country"`: National aggregate
- `"region"`: Brazilian geographic regions (5 regions)
- `"state"`: State-level data (27 units)
- `"municipality"`: All 5,570+ municipalities

```r
geo_level = "state"  # character string
```

### 4. **time_period**

Specifies which year(s) to download.

```r
time_period = 2020              # single year
time_period = c(2010, 2020)     # specific years
time_period = 2010:2020         # range of years
```

**Note**: All datasets available from 1974 onwards, though aquaculture more complete from 2000s.

### 5. **language**

Output language for variable names.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

***

## Examples

### Example 1: Livestock Specialization by State

Identify regional specialization in different livestock:

```{r eval=FALSE}
library(dplyr)

# Load livestock inventory for 2020
livestock <- load_ppm(
  dataset = "ppm_livestock_inventory",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2020,
  language = "eng"
)

# Livestock composition by state
livestock_composition <- livestock %>%
  group_by(state) %>%
  mutate(
    total_animals = sum(number_of_animals, na.rm = TRUE)
  ) %>%
  group_by(state, animal_species) %>%
  summarize(
    num_animals = sum(number_of_animals, na.rm = TRUE),
    pct_of_state = (num_animals / first(total_animals)) * 100,
    .groups = 'drop'
  ) %>%
  pivot_wider(
    names_from = animal_species,
    values_from = num_animals,
    values_fill = 0
  )

# States with largest cattle herds
cattle_states <- livestock %>%
  filter(animal_species == "Cattle") %>%
  group_by(state) %>%
  summarize(cattle_herd = sum(number_of_animals, na.rm = TRUE), .groups = 'drop') %>%
  arrange(desc(cattle_herd)) %>%
  head(15)

print(cattle_states)
```

### Example 2: Dairy Productivity Analysis

Compare dairy cow productivity across regions:

```{r eval=FALSE}
library(dplyr)

# Load dairy cow data
dairy <- load_ppm(
  dataset = "ppm_cow_farming",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2020,
  language = "eng"
)

# Dairy productivity metrics
dairy_analysis <- dairy %>%
  mutate(
    productivity_liter_per_cow = milk_production_liters / milked_cows
  ) %>%
  group_by(state) %>%
  summarize(
    total_milked_cows = sum(milked_cows, na.rm = TRUE),
    total_milk_production = sum(milk_production_liters, na.rm = TRUE),
    avg_productivity = mean(productivity_liter_per_cow, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_milk_production))

print("Milk production by state:")
print(dairy_analysis)

# Identify most and least productive states
most_productive <- dairy_analysis %>%
  arrange(desc(avg_productivity)) %>%
  head(10)

print("Most productive dairy states:")
print(most_productive)
```

### Example 3: Animal Product Economics

Analyze the economic value of different animal products:

```{r eval=FALSE}
library(dplyr)

# Load animal product data
animal_products <- load_ppm(
  dataset = "ppm_animal_origin_production",
  raw_data = FALSE,
  geo_level = "country",
  time_period = 2020,
  language = "eng"
)

# Product values
product_value <- animal_products %>%
  group_by(product_type) %>%
  summarize(
    total_production = sum(quantity_produced, na.rm = TRUE),
    total_value = sum(production_value, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    value_per_unit = total_value / total_production
  ) %>%
  arrange(desc(total_value))

print(product_value)

# Milk dominance in animal products
print("Economic importance of each product:")
product_value %>%
  mutate(
    pct_of_total = (total_value / sum(total_value)) * 100
  ) %>%
  arrange(desc(pct_of_total))
```

### Example 4: Regional Livestock Structure

Compare livestock sectors across regions:

```{r eval=FALSE}
library(dplyr)

# Load regional data
livestock <- load_ppm(
  dataset = "ppm_livestock_inventory",
  raw_data = FALSE,
  geo_level = "region",
  time_period = 2020,
  language = "eng"
)

# Regional livestock profile
regional_livestock <- livestock %>%
  group_by(region, animal_species) %>%
  summarize(
    num_animals = sum(number_of_animals, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  pivot_wider(
    names_from = animal_species,
    values_from = num_animals,
    values_fill = 0
  )

print(regional_livestock)

# Specialization indices
# (Which region has highest proportion of each animal type)
specialization <- livestock %>%
  group_by(animal_species) %>%
  mutate(
    national_total = sum(number_of_animals, na.rm = TRUE)
  ) %>%
  group_by(region, animal_species) %>%
  summarize(
    regional_share = sum(number_of_animals, na.rm = TRUE) / first(national_total) * 100,
    .groups = 'drop'
  )

print(specialization)
```


### Example 5: Wool Production Analysis

Analyze sheep farming and wool production:

```{r eval=FALSE}
library(dplyr)

# Load sheep farming data
sheep <- load_ppm(
  dataset = "ppm_sheep_farming",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2020,
  language = "eng"
)

# Wool production by state
wool_production <- sheep %>%
  group_by(state) %>%
  summarize(
    total_sheep = sum(total_sheep, na.rm = TRUE),
    sheared_sheep = sum(sheared_sheep, na.rm = TRUE),
    total_fleece_weight = sum(fleece_weight_kg, na.rm = TRUE),
    avg_fleece_per_sheep = mean(fleece_weight_kg / sheared_sheep, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_fleece_weight))

print(wool_production)

# Shearing percentage
wool_production %>%
  mutate(
    shearing_percentage = (sheared_sheep / total_sheep) * 100
  ) %>%
  select(state, total_sheep, shearing_percentage)
```

### Example 6: Aquaculture Growth

Analyze emerging aquaculture sector:

```{r eval=FALSE}
library(dplyr)

# Load aquaculture data for recent years
aquaculture <- load_ppm(
  dataset = "ppm_aquaculture",
  raw_data = FALSE,
  geo_level = "state",
  time_period = 2015:2020,
  language = "eng"
)

# Aquaculture growth by type
aquaculture_by_type <- aquaculture %>%
  group_by(year, aquaculture_type) %>%
  summarize(
    production = sum(quantity_produced, na.rm = TRUE),
    value = sum(production_value, na.rm = TRUE),
    .groups = 'drop'
  )

# Leading states in aquaculture
leading_aquaculture <- aquaculture %>%
  filter(year == 2020) %>%
  group_by(state) %>%
  summarize(
    total_production = sum(quantity_produced, na.rm = TRUE),
    total_value = sum(production_value, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_value)) %>%
  head(15)

print("Top aquaculture states:")
print(leading_aquaculture)
```

***

## Data Notes

### Data Structure

Each record typically contains:
- Geographic identifier (state or municipality)
- Year
- Animal species or product type
- Quantity (number of animals or production volume)
- Value (if applicable)
- Number of establishments

### Units of Measurement

- **Livestock counts**: Number of animals
- **Milk**: Liters
- **Eggs**: Dozens or units (verify in data)
- **Honey**: Kilograms
- **Wool**: Kilograms
- **Aquaculture**: Kilograms or tons

### Raw vs. Treated Data

- **Raw data**: IBGE original format, Portuguese names
- **Treated data**: English variable names, standardized units

### Important Limitations

1. **Survey-based data**: Subject to sampling and reporting error
2. **Informal operations**: May undercount small or informal livestock operations
3. **Data lag**: Published with delay; recent years may not be available
4. **Aquaculture newer**: Aquaculture data less complete for very early years
5. **Methodology changes**: Survey methods may evolve; can affect comparability

***
