---
title: "PRODES"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRODES}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

[PRODES](https://www.gov.br/inpe/) (Projeto de Monitoramento da Floresta Amazônica Brasileira - Project for Monitoring the Brazilian Amazon Forest) is Brazil's primary satellite-based deforestation monitoring system operated by [INPE](https://www.inpe.br/) (National Institute for Space Research).

This dataset provides:

- **Amazon deforestation monitoring**: Official deforestation data using satellite imagery
- **Multiple landscape classes**: Deforestation, residual deforestation, native vegetation, hydrography, non-forest areas, clouds
- **Raster data**: High-resolution spatial data covering entire Legal Amazon
- **Municipality aggregates**: Affected areas per municipality for non-spatial analysis
- **Cumulative and incremental data**: Total deforestation since 1988 and year-by-year changes
- **Long time series**: Available from 2007 onwards (cumulative) and incrementally by year
- **Official baseline**: Used by Brazilian government for forest policy and enforcement

PRODES is the authoritative source for Amazon deforestation statistics, used for environmental monitoring, policy evaluation, and international reporting on Brazilian forest protection.

### Data Source and Methodology

PRODES monitoring:
- Uses satellite imagery from multiple sources (Landsat, CBERS, others)
- Automated and manual analysis to detect clear-cut deforestation
- Annual assessment of forest loss
- Recently (2020+) raster data published through [TerraBrasilis](https://terrabrasilis.dpi.inpe.br/)
- Data available as both raster files and aggregated by municipality

For more information, visit [INPE PRODES Project](https://www.gov.br/inpe/) and [TerraBrasilis](https://terrabrasilis.dpi.inpe.br/).

---

## Available Datasets

### **1. deforestation**

Clear-cut deforestation areas (complete forest loss).

- **Coverage**: Entire Legal Amazon region
- **Time period**: 
  - 2007: Cumulative deforestation from 1988 to 2007
  - 2008-2023: Annual incremental deforestation (year-specific)
- **Data types**: 
  - Raw: Raster/grid data (SpatRaster format)
  - Treated: Municipality-level area aggregates
- **Variables**: Deforested area (hectares) by municipality
- **Use cases**:
  - Track annual deforestation trends
  - Identify deforestation hotspots
  - Cumulative forest loss assessment
  - Municipality-level deforestation analysis

### **2. residual_deforestation**

Deforestation that was not captured in previous surveys (detected through improved methodology).

- **Coverage**: Legal Amazon
- **Time period**: 2010-2023
- **Data types**: Raster or municipality aggregates
- **Variables**: Additional detected deforestation (hectares)
- **Use cases**:
  - More comprehensive deforestation assessment
  - Correct for historical detection gaps
  - Full accounting of forest loss

### **3. native_vegetation**

Remaining native forest and natural vegetation areas.

- **Coverage**: All vegetation in Legal Amazon
- **Time period**: 2023 (snapshot of current state)
- **Data types**: Raster only
- **Variables**: Forest and vegetation extent
- **Use cases**:
  - Assess remaining forest cover
  - Identify conservation priority areas
  - Vegetation extent mapping

### **4. hydrography**

Water bodies and hydrographic features.

- **Coverage**: All water features in Amazon
- **Time period**: 2023
- **Data types**: Raster only
- **Variables**: Water body extent
- **Use cases**:
  - Environmental impact assessment
  - Hydrology and water resource mapping

### **5. non_forest**

Non-forest areas including savanna, grasslands, and other vegetation types.

- **Coverage**: Legal Amazon region
- **Time period**: 2023
- **Data types**: Raster only
- **Variables**: Non-forest vegetation extent
- **Use cases**:
  - Distinguish non-forest from forest areas
  - Natural vegetation assessment
  - Land-use classification

### **6. clouds**

Cloud cover in satellite imagery (data quality indicator).

- **Coverage**: Legal Amazon
- **Time period**: 2023
- **Data types**: Raster only
- **Variables**: Cloud-covered areas where analysis is uncertain
- **Use cases**:
  - Assess data quality and coverage
  - Identify areas needing re-analysis
  - Quantify monitoring gaps

---

## Important Data Characteristics

### Raw vs. Treated Data

- **Raw data** (`raw_data = TRUE`): Returns SpatRaster objects (raster grids) from TerraBrasilis
- **Treated data** (`raw_data = FALSE`): Aggregated to municipality level showing total affected area
- **Raster resolution**: Typically ~30-meter pixels
- **Large size**: Raster data is high-resolution and may be large

### Cumulative vs. Incremental Data

| Year | Type | Definition |
|------|------|-----------|
| 2007 | Cumulative | All deforestation from 1988-2007 |
| 2008-2023 | Incremental | Deforestation detected in that specific year |

When analyzing time trends, be aware that 2007 includes 19 years of accumulated loss.

### Data Organization

- **Raster format**: High-resolution grid with classified pixels (deforestation, forest, etc.)
- **Municipality level**: Treated data aggregates raster to municipality boundaries
- **Non-spatial**: Municipality aggregates don't retain spatial location (summarized only)

---

## Function Parameters

### 1. **dataset**

Selects which landscape classification to download.

```r
dataset = "deforestation"             # Clear-cut deforestation (main product)
dataset = "residual_deforestation"    # Previously undetected deforestation
dataset = "native_vegetation"         # Remaining forests and vegetation (2023 only)
dataset = "hydrography"               # Water bodies (2023 only)
dataset = "non_forest"                # Savanna and other non-forest (2023 only)
dataset = "clouds"                    # Cloud cover (2023 only)
```

### 2. **raw_data**

Controls data format: raster or municipality aggregates.

- `TRUE`: Returns raw raster data (SpatRaster objects)
- `FALSE`: Returns municipality-level aggregated area data

```r
raw_data = FALSE  # logical
```

**Important**: Raster data is large; ensure sufficient storage and memory.

### 3. **time_period**

Specifies which year(s) to download.

**Available by dataset:**

| Dataset | Available Years |
|---------|---|
| `deforestation` | 2007 (cumulative), 2008-2023 (incremental) |
| `residual_deforestation` | 2010-2023 |
| `native_vegetation` | 2023 only |
| `hydrography` | 2023 only |
| `non_forest` | 2023 only |
| `clouds` | 2023 only |

```r
time_period = 2020              # single year
time_period = c(2015, 2020)     # multiple years
time_period = 2015:2020         # range of years (for deforestation only)
```

### 4. **language**

Output language.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

---

## Examples


### Example 1: Deforestation Hotspots

Identify municipalities with highest deforestation:

```{r eval=FALSE}
library(dplyr)

# Load recent deforestation data
deforestation <- load_prodes(
  dataset = "deforestation",
  raw_data = FALSE,
  time_period = 2023,
  language = "eng"
)

# Municipalities with most recent deforestation
hotspots <- deforestation %>%
  arrange(desc(deforested_area_hectares)) %>%
  select(municipality, state, year, deforested_area_hectares) %>%
  head(30)

print("Deforestation hotspots (2023):")
print(hotspots)

# Cumulative deforestation from 2008-2023
deforestation_series <- load_prodes(
  dataset = "deforestation",
  raw_data = FALSE,
  time_period = 2008:2023,
  language = "eng"
)

total_by_municipality <- deforestation_series %>%
  group_by(municipality, state) %>%
  summarize(
    cumulative_deforestation = sum(deforested_area_hectares, na.rm = TRUE),
    years_with_activity = n_distinct(year),
    .groups = 'drop'
  ) %>%
  arrange(desc(cumulative_deforestation)) %>%
  head(30)

print("Municipalities with most cumulative deforestation (2008-2023):")
print(total_by_municipality)
```

### Example 2: Regional Deforestation Patterns

Compare deforestation across states and regions:

```{r eval=FALSE}
library(dplyr)

deforestation <- load_prodes(
  dataset = "deforestation",
  raw_data = FALSE,
  time_period = 2023,
  language = "eng"
)

# State-level deforestation
state_deforest <- deforestation %>%
  group_by(state) %>%
  summarize(
    total_deforestation = sum(deforested_area_hectares, na.rm = TRUE),
    num_municipalities_affected = n_distinct(municipality),
    avg_per_municipality = mean(deforested_area_hectares, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_deforestation))

print(state_deforest)

# Share of total deforestation
state_deforest <- state_deforest %>%
  mutate(
    pct_of_total = (total_deforestation / sum(total_deforestation)) * 100
  )

# Top states account for what %?
top5 <- state_deforest %>%
  head(5) %>%
  summarize(top5_pct = sum(pct_of_total))

print(paste("Top 5 states account for", round(top5$top5_pct, 1), "% of 2023 deforestation"))
```

### Example 3: Cumulative Forest Loss

Analyze total deforestation since 2007:

```{r eval=FALSE}
library(dplyr)

# Load 2007 (cumulative from 1988-2007) plus all years 2008-2023
all_deforestation <- load_prodes(
  dataset = "deforestation",
  raw_data = FALSE,
  time_period = c(2007, 2008:2023),
  language = "eng"
)

# Cumulative loss by 2023
cumulative_2023 <- all_deforestation %>%
  group_by(municipality, state) %>%
  summarize(
    total_deforestation_to_2023 = sum(deforested_area_hectares, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_deforestation_to_2023))

print("Municipalities with most cumulative deforestation (1988-2023):")
print(head(cumulative_2023, 20))

# Total Amazon deforestation
total_amazon_loss <- sum(cumulative_2023$total_deforestation_to_2023, na.rm = TRUE)
print(paste("Total Amazon deforestation through 2023:", 
            round(total_amazon_loss / 1000000, 1), "million hectares"))
```
---

## Data Notes

### Raw Raster Data Characteristics

- **Format**: SpatRaster (from terra package)
- **Resolution**: ~30 meter pixels
- **Coverage**: Complete Legal Amazon region
- **Size**: Very large files; may require significant storage/memory
- **Classification**: Pixel values indicate land cover class (deforestation, forest, etc.)
- **CRS**: UTM zones or standard projection; verify after loading

### Municipality Aggregates

- **Simplified for analysis**: Loses spatial detail but much smaller files
- **Aggregation**: Total deforested area per municipality per year
- **Ideal for**: Time series, trend analysis, comparative studies

### Data Quality and Accuracy

1. **Detection accuracy**: ~95% for clear-cut deforestation
2. **Minimum mapping unit**: Small forest clearances may not be detected
3. **Cloud contamination**: Clouds may prevent detection in some areas (see clouds dataset)
4. **Seasonal effects**: Deforestation easier to detect in dry season

### Limitations

1. **Clear-cut only**: Degradation and partial logging not captured (see DEGRAD for that)
2. **Legal Amazon only**: Data limited to Legal Amazon definition; doesn't cover all forest regions
3. **Raster data very large**: May require specialized tools and significant computing resources
4. **Historical changes**: INPE occasionally revises historical estimates as methodology improves
5. **Recent data provisional**: 2023 data may be subject to revision as final processing completes

---

