---
title: "PRODES"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRODES}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The PRODES project uses satellites to monitor deforestation in Brazil's Legal Amazon. The raw data reports total and incremental (year-by-year) low-cut deforested area at the municipality level, going back to the year 2000.

Data is collected based on the PRODES-year, which starts at August 1st and ends on July 31st. Accordingly, 2018 deforestation data covers the period from 01/08/2017 to 31/07/2018.

INPE's most recent data is now published at [TerraBrasilis](http://terrabrasilis.dpi.inpe.br/downloads/). We have refrained from updating to this new source, as it only contains detailed spatial data, rather than agregated, municipality-level data.

***
**Options:**

  1. **dataset**: `"deforestation"`

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. 

  3. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
***

**Examples:**

```{r,eval=FALSE}
# Download treated data (raw_data = FALSE)
# in portuguese (language = 'pt').
data <- load_prodes(
  raw_data = FALSE,
  language = "pt"
)
```

