---
title: "PRODES"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRODES}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The PRODES project uses satellites to monitor deforestation in Brazil's Legal Amazon. The raw data reports total and incremental (year-by-year) low-cut deforested area at the municipality level.

The data made available in this package goes back to the year 2000, with ongoing updates. In line with INPE's API, requesting data for an unavailable year does not yield an error, but rather a best effort response (columns regarding observation data are filled with default values).

Data is collected based on the PRODES-year, which starts at August 1st and ends on July 31st. Accordingly, 2018 deforestation data covers the period from 01/08/2017 to 31/07/2018. 

```
There are four parameters in this function:
  
  1. dataset = "prodes"
 
  2. raw_data: there are two options:
  # TRUE: if you want the raw data.
  # FALSE: if you want the processed version of the data. 
 
  3. time_period: available from 2000 to 2020
  
  4. language: you can choose between portuguese and english
```

```{r,eval=FALSE}
# Download treated data (raw_data = FALSE) from 2010 (time_period = 2010) 
# in portuguese (language = 'pt').
data <- load_prodes(dataset = "prodes", 
                    raw_data = FALSE,
                    time_period = 2010,
                    language = 'pt')  
```

