---
title: "PRODES"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRODES}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The PRODES project uses satellites to monitor deforestation in Brazil's Legal Amazon. The raw data reports total and incremental (year-by-year) low-cut deforested area at the municipality level, going back to the year 2007.

INPE's most recent data is now published at [TerraBrasilis](http://terrabrasilis.dpi.inpe.br/downloads/). We read their full [raster data](https://terrabrasilis.dpi.inpe.br/geonetwork/srv/eng/catalog.search#/metadata/507294db-a789-42dd-9158-9dea77d9293f) for the Legal Amazon region and extract values onto the map of Brazilian municipalities.

***
**Options:**

  1. **dataset**: `"deforestation"`, `"residual_deforestation"`, `"native_vegetation"`, `"hydrography"`, `"non_forest"`, or `"clouds"`

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally, read as a SpatRaster.
      * `FALSE`: if you want the treated version of the data, measuring affected areas per municipality.
  
  3. **time_period**: picks the years for which the data will be downloaded, under the following constraints:
      * For dataset `"deforestation"`, it can be between 2007 and 2023. Deforestation for 2007 includes all cumulative deforestation up to 2007. For other years, deforestation is incremental;
      * For dataset `"residual_deforestation"`, it can be between 2010 and 2023;
      * For all other datasets, only the year 2023 is available.
  
  4. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
***

**Examples:**

```{r,eval=FALSE}
# Download treated data (raw_data = FALSE)
# in portuguese (language = 'pt').
data <- load_prodes(
  dataset = "deforestation",
  raw_data = FALSE,
  time_period = 2020:2023,
  language = "pt"
)
```

