---
title: "SEEG"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SEEG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview

SEEG (Sistema de Estimativa de Emissões e Remoções de Gases de Efeito Estufa - System of Estimates of Emissions and Removals of Greenhouse Gases) is Brazil's most comprehensive greenhouse gas emissions database developed by [Observatório do Clima](https://www.observatoriodoclima.org.br/) (Climate Observatory).

This dataset provides:

- **Greenhouse gas emissions**: Complete estimates of all major climate-relevant gases
- **Multi-sector coverage**: Agriculture, energy, land use, industry, waste
- **Sub-sectoral detail**: Detailed breakdowns within each sector
- **Municipality and state levels**: Geographic disaggregation for regional analysis
- **Time series**: Historical data from 2000 onwards
- **Removal accounting**: Also includes carbon sequestration and removals
- **Comprehensive methodology**: Based on Brazilian national inventory standards
- **Transparent assumptions**: Well-documented methodology and data sources

SEEG is the primary tool for understanding Brazil's greenhouse gas emissions profile, tracking progress toward climate goals, identifying emission hotspots, and supporting climate policy.

### Data Source and Methodology

SEEG emissions estimates are compiled using:
- Government data from multiple agencies (MAPA, IBGE, ANP, etc.)
- Satellite monitoring of deforestation and land use
- International IPCC methodology standards
- Peer-reviewed scientific research
- Regular updates as new government data becomes available

For more information, visit [SEEG Project](https://seeg.org.br/) and [Observatório do Clima](https://www.observatoriodoclima.org.br/).

***

## Available Datasets

### **1. seeg (All Sectors Combined)**

Complete greenhouse gas emissions across all sectors in one dataset.

- **Coverage**: All emission sources in Brazil
- **Sectors included**: All five (agriculture, energy, land use, industry, waste)
- **Time period**: 2000-2018
- **Geographic levels**: Country, State, Municipality
- **Key variables**: Total emissions (CO₂e), by sector and sub-sector
- **Format**: Comprehensive view of Brazil's total emissions profile
- **Note**: Only available with `raw_data = TRUE`
- **Use cases**:
  - Understand overall emissions landscape
  - Identify dominant emission sources
  - Track total emissions trends over time

### **2. seeg_farming (Agricultural and Livestock Emissions)**

Greenhouse gas emissions from agriculture and livestock activities.

- **Coverage**: All agricultural and livestock production
- **Time period**: 2000-2018
- **Geographic levels**: Country, State, Municipality
- **Key variables**: Emissions from cattle, crop production, soil management, manure
- **Dominant source**: Usually the largest single emissions sector in Brazil
- **Components**: 
  - Livestock (enteric fermentation, manure)
  - Crop production and soil management
  - Agricultural land preparation
- **Use cases**:
  - Assess agricultural emission contributions
  - Identify highest-emission municipalities
  - Evaluate livestock and farming intensity
  - Policy targets for agricultural emissions reduction

### **3. seeg_energy (Energy Sector Emissions)**

Emissions from energy production and consumption.

- **Coverage**: All energy-related emissions
- **Time period**: 2000-2018
- **Geographic levels**: Country, State, Municipality
- **Key variables**: Emissions from electricity, transport, heating, fuel production
- **Components**:
  - Energy generation and distribution
  - Transportation fuels
  - Energy consumption
  - Industrial energy use
- **Use cases**:
  - Understand energy sector contribution to climate change
  - Track renewable vs. fossil fuel impacts
  - Identify regional energy emission patterns

### **4. seeg_land (Land Use Change Emissions)**

Emissions and removals from changes in forest cover and land use.

- **Coverage**: Deforestation, forest degradation, reforestation effects
- **Time period**: 2000-2018
- **Geographic levels**: Country, State, Municipality
- **Key variables**: Net emissions/removals from land use change
- **Components**:
  - Deforestation and forest loss
  - Forest degradation
  - Reforestation and afforestation
  - Vegetation conversion
- **Importance**: Often largest single contributor to Brazil's emissions
- **Use cases**:
  - Analyze deforestation climate impact
  - Identify reforestation opportunities
  - Assess forest conservation value
  - Link with PRODES and DETER deforestation data

### **5. seeg_industry (Industrial Process Emissions)**

Emissions from manufacturing and industrial processes.

- **Coverage**: All industrial sectors
- **Time period**: 2000-2018
- **Geographic levels**: Country, State, Municipality
- **Key variables**: Emissions from cement, chemicals, metals, minerals, other manufacturing
- **Components**:
  - Chemical production (ammonia, soda ash, etc.)
  - Metal production (iron, aluminum, others)
  - Mineral processing (cement, lime, glass)
  - Other industrial processes
- **Use cases**:
  - Identify industrial emission hotspots
  - Regional manufacturing impacts
  - Process-specific emission reduction opportunities

### **6. seeg_residuals (Waste and Residuals Emissions)**

Emissions from waste management, landfills, and waste treatment.

- **Coverage**: All waste-related emissions
- **Time period**: 2000-2018
- **Geographic levels**: Country, State, Municipality
- **Key variables**: Emissions from solid waste, wastewater treatment, waste treatment
- **Components**:
  - Landfill methane emissions
  - Wastewater treatment
  - Waste disposal and treatment
  - Municipal solid waste management
- **Use cases**:
  - Assess waste sector contributions
  - Identify waste management improvement opportunities
  - Evaluate circular economy potential

***

## Important Data Characteristics

### Collection 9 Data

The data provided is from SEEG's Collection 9:
- **Time period**: 2000-2018
- **Methodology**: Latest available when data was compiled
- **Quality**: Peer-reviewed and validated
- **Revisions**: May be updated in future SEEG collections as better data becomes available

### Emissions Units

- **Standard unit**: Gigatonnes CO₂ equivalent (Gt CO₂e)
- **CO₂e equivalence**: Uses global warming potentials (GWP) to convert CH₄ and N₂O to CO₂ equivalent
- **Consistency**: Allows comparison across different gases and sectors

### Download Considerations

 **Important**: The complete SEEG dataset is quite large. When downloading:
- Entire datasets are downloaded as single files; year selection is limited
- A stable, high-speed internet connection is recommended
- Downloads may take time depending on connection speed
- Ensure sufficient disk space for storage

***

## Function Parameters

### 1. **dataset**

Selects which emission sector(s) to download.

```r
dataset = "seeg"              # All sectors (raw_data = TRUE only)
dataset = "seeg_farming"      # Agriculture and livestock
dataset = "seeg_energy"       # Energy sector
dataset = "seeg_land"         # Land use changes
dataset = "seeg_industry"     # Industrial processes
dataset = "seeg_residuals"    # Waste and residuals
```

### 2. **raw_data**

Controls whether to download original or processed data.

- `TRUE`: Returns raw SEEG data format (more detailed)
- `FALSE`: Returns treated data with English variable names and standardized format

```r
raw_data = FALSE  # logical
```

### 3. **geo_level**

Specifies geographic aggregation level.

- `"country"`: National total
- `"state"`: State-level emissions (27 units)
- `"municipality"`: All 5,570+ municipalities

```r
geo_level = "state"  # character string
```

### 4. **language**

Output language for variable names and labels.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

**Note on timing**: Downloads may take considerable time due to file size.

***
## Examples

### Example 1: Total National Emissions by Sector

Understand Brazil's overall emissions profile:

```{r eval=FALSE}
library(dplyr)
library(ggplot2)

# Load all sectors (raw data format)
all_emissions <- load_seeg(
  dataset = "seeg",
  raw_data = TRUE,
  geo_level = "country",
  language = "eng"
)

# Total emissions by sector for latest available year
sector_breakdown <- all_emissions %>%
  group_by(sector) %>%
  summarize(
    total_emissions = sum(emissions_co2e, na.rm = TRUE),
    pct_of_total = (total_emissions / sum(all_emissions$emissions_co2e, na.rm = TRUE)) * 100,
    .groups = 'drop'
  ) %>%
  arrange(desc(total_emissions))

print("Brazil's emissions by sector:")
print(sector_breakdown)

# Visualize sector contributions
ggplot(sector_breakdown, aes(x = reorder(sector, total_emissions), y = total_emissions)) +
  geom_col(fill = "darkred") +
  coord_flip() +
  labs(title = "Brazil's Greenhouse Gas Emissions by Sector",
       x = "Sector",
       y = "Emissions (CO₂e)") +
  theme_minimal()
```

### Example 2: Agricultural Emissions Analysis

Analyze the largest emissions sector for Brazil:

```{r eval=FALSE}
library(dplyr)

# Load agricultural emissions by state
farming <- load_seeg(
  dataset = "seeg_farming",
  raw_data = FALSE,
  geo_level = "state",
  language = "eng"
)

# States with highest agricultural emissions
top_farm_states <- farming %>%
  group_by(state) %>%
  summarize(
    total_ag_emissions = sum(emissions_co2e, na.rm = TRUE),
    num_sources = n_distinct(source),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_ag_emissions)) %>%
  head(15)

print("Top agricultural emission states:")
print(top_farm_states)

# Breakdown of agricultural sources (livestock vs crop)
ag_sources <- farming %>%
  group_by(source) %>%
  summarize(
    emissions = sum(emissions_co2e, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(emissions))

print("Agricultural emission sources:")
print(ag_sources)
```

### Example 3: Land Use Change Emissions

Understand the climate impact of deforestation:

```{r eval=FALSE}
library(dplyr)

# Load land use change emissions
land_use <- load_seeg(
  dataset = "seeg_land",
  raw_data = FALSE,
  geo_level = "state",
  language = "eng"
)

# Net emissions/removals by state
state_net_emissions <- land_use %>%
  group_by(state) %>%
  summarize(
    net_emissions = sum(emissions_co2e, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(net_emissions))

print("Land use change emissions by state (positive = emissions from deforestation):")
print(state_net_emissions)

# States with highest deforestation impact
high_emission_states <- state_net_emissions %>%
  filter(net_emissions > 0) %>%
  top_n(10, net_emissions)

print("States with highest deforestation emissions:")
print(high_emission_states)
```

### Example 4: Emissions Trends Over Time

Track how emissions have changed from 2000 to 2018:

```{r eval=FALSE}
library(dplyr)
library(ggplot2)

# Load all sectors for country level
all_sectors <- load_seeg(
  dataset = "seeg",
  raw_data = TRUE,
  geo_level = "country",
  language = "eng"
)

# Annual emissions trends
annual_emissions <- all_sectors %>%
  group_by(year) %>%
  summarize(
    total_emissions = sum(emissions_co2e, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(year)

print(annual_emissions)

# Visualize trends
ggplot(annual_emissions, aes(x = year, y = total_emissions)) +
  geom_line(size = 1, color = "darkred") +
  geom_point(size = 2) +
  labs(title = "Brazil's Total Greenhouse Gas Emissions (2000-2018)",
       x = "Year",
       y = "Total Emissions (CO₂e)") +
  theme_minimal()

# Calculate trend
first_year <- annual_emissions$total_emissions[1]
last_year <- annual_emissions$total_emissions[nrow(annual_emissions)]
change_pct <- ((last_year - first_year) / first_year) * 100

print(paste("Total change 2000-2018:", round(change_pct, 1), "%"))
```

### Example 5: Municipal Emission Hotspots

Identify highest-emitting municipalities:

```{r eval=FALSE}
library(dplyr)

# Load municipality-level data
all_sectors_munic <- load_seeg(
  dataset = "seeg",
  raw_data = TRUE,
  geo_level = "municipality",
  language = "eng"
)

# Highest-emitting municipalities
top_municipalities <- all_sectors_munic %>%
  group_by(municipality, state) %>%
  summarize(
    total_emissions = sum(emissions_co2e, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_emissions)) %>%
  head(30)

print("Highest-emitting municipalities in Brazil:")
print(top_municipalities)

# Emission concentration: top municipalities account for what %?
total_emissions <- sum(all_sectors_munic$emissions_co2e, na.rm = TRUE)
top100 <- top_municipalities %>%
  head(100) %>%
  summarize(top100_total = sum(total_emissions))

top100_pct <- (top100$top100_total / total_emissions) * 100
print(paste("Top 100 municipalities account for", round(top100_pct, 1), "% of emissions"))
```

### Example 6: Sector Comparison by State

Compare sectoral composition across states:

```{r eval=FALSE}
library(dplyr)
library(tidyr)

# Load individual sector datasets
farming <- load_seeg(dataset = "seeg_farming", raw_data = FALSE, 
                     geo_level = "state", language = "eng") %>%
  mutate(sector = "Agriculture") %>%
  group_by(state, sector) %>%
  summarize(emissions = sum(emissions_co2e, na.rm = TRUE), .groups = 'drop')

energy <- load_seeg(dataset = "seeg_energy", raw_data = FALSE,
                    geo_level = "state", language = "eng") %>%
  mutate(sector = "Energy") %>%
  group_by(state, sector) %>%
  summarize(emissions = sum(emissions_co2e, na.rm = TRUE), .groups = 'drop')

land_use <- load_seeg(dataset = "seeg_land", raw_data = FALSE,
                      geo_level = "state", language = "eng") %>%
  mutate(sector = "Land Use") %>%
  group_by(state, sector) %>%
  summarize(emissions = sum(emissions_co2e, na.rm = TRUE), .groups = 'drop')

# Combine
combined <- bind_rows(farming, energy, land_use)

# State profiles
state_profile <- combined %>%
  pivot_wider(
    names_from = sector,
    values_from = emissions,
    values_fill = 0
  ) %>%
  mutate(
    total = Agriculture + Energy + `Land Use`,
    ag_share = (Agriculture / total) * 100,
    energy_share = (Energy / total) * 100,
    land_share = (`Land Use` / total) * 100
  )

print(state_profile)
```
***

## Data Notes

### Emission Sources Included

SEEG includes all major anthropogenic emission sources:
- Agriculture (livestock, crops, soil)
- Energy (electricity, transport, heating)
- Land use change (deforestation, afforestation)
- Industrial processes (cement, chemicals, metals)
- Waste (landfills, wastewater)

### Methodology

Estimates follow:
- IPCC guidelines for national greenhouse gas inventories
- Brazilian national inventory standards
- International best practices
- Transparent, documented assumptions

### Data Quality

- Peer-reviewed methodology
- Validated against government data
- Uncertainty ranges available in detailed products
- Regular methodology updates

### Limitations

1. **Fixed time period**: Collection 9 covers 2000-2018 only
2. **File size**: Large downloads; requires good internet
3. **Year aggregation**: Cannot select individual years; entire dataset downloaded
4. **Revisions**: Methodology may change in future SEEG releases
5. **Sub-national uncertainty**: Municipal and state estimates have higher uncertainty than national

***
