---
title: "SEEG"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SEEG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview

SEEG (Sistema de Estimativa de Emissões e Remoções de Gases de Efeito Estufa - System of Estimates of Emissions and Removals of Greenhouse Gases) is Brazil's most comprehensive greenhouse gas emissions database, developed by [Observatório do Clima](https://oc.eco.br/) (Climate Observatory).

This dataset provides:

- **Greenhouse gas emissions**: estimates covering emissions and removals across all major sectors
- **Multi-sector coverage**: agriculture, energy, land use change, industrial processes, waste
- **Municipality, state, and country levels**: geographic disaggregation for regional analysis
- **Time series**: 1970-2024

For more information, visit [SEEG Project](https://www.seeg.org.br/) and [Observatório do Clima](https://oc.eco.br/).

***

## Data Version

The package reads SEEG's **v13.0** municipal summary workbook (`seeg.eco.br`). Categories are
organized as `Setor de emissão` / `Categoria emissora` / `Sub-categoria emissora` / `Recorte` /
`Atividade geral`. There is no `time_period` argument -- every call returns the full series.
Exact coverage always tracks the manifest (`datasets_link(source = "seeg")` or
`inst/extdata/manifest/v1/datasets_link.csv`); check there before assuming this has changed.

### Detail not included

A few dimensions are not available in the current data:

- No `gas`/`produto` breakdown -- emissions are not split by individual gas.
- **seeg_farming**: no species/crop-level detail.
- **seeg_industry**: no input/technology-level detail (partly offset by an actual/potential
  emission split and a finer end-use breakdown).
- **seeg_energy**: no fuel-type breakdown (gasoline, diesel, LPG, etc.).
- **seeg_land**: no **biome** dimension and no origin-to-destination **land-cover transition
  matrix** -- only four coarser categorical fields.

***

## Available Datasets

### **1. seeg (All Sectors Combined)**

Complete greenhouse gas emissions across all sectors in one dataset. **Only available with
`raw_data = TRUE`** -- it cannot be requested treated (`raw_data = FALSE`).

- **Coverage**: all sectors, 1970-2024
- **Geographic levels**: country, state, municipality
- **Format**: the cleaned source sheet -- one row per municipality x category path, with one
  numeric column per year (`x1970`...`x2024`)

### **2. seeg_farming (Agricultural and Livestock Emissions)**

Sector: `Agropecuaria`. **Only available with `raw_data = FALSE`.**

### **3. seeg_energy (Energy Sector Emissions)**

Sector: `Energia`. **Only available with `raw_data = FALSE`.** No fuel-type breakdown (see
"Detail not included" above).

### **4. seeg_land (Land Use Change Emissions)**

Sector: `Mudanca de Uso da Terra e Floresta`. **Only available with `raw_data = FALSE`.** No
biome dimension and no land-cover transition matrix (see "Detail not included" above).

### **5. seeg_industry (Industrial Process Emissions)**

Sector: `Processos Industriais`. **Only available with `raw_data = FALSE`.**

### **6. seeg_residuals (Waste and Residuals Emissions)**

Sector: `Residuos`. **Only available with `raw_data = FALSE`.**

***

## Output Columns (`raw_data = FALSE`)

Every `seeg_*` sector dataset returns long-format data, one row per category combination and
year, with a single `value` (eng) / `Valor` (pt) column. The columns available depend on
`geo_level` and on which 4th categorical dimension that sector carries:

| geo_level | eng columns | pt columns |
|---|---|---|
| municipality | year, city, state, ibge, sector, emission_category, emission_subcategory, **&lt;4th dim&gt;**, general_activity, emission_type, value | Ano, municipio, estado, ibge, setor, categoria_emissao, subcategoria_emissao, **&lt;4th dim&gt;**, atividade_geral, tipo_emissao, Valor |
| state | (drops city, ibge) | (drops municipio, ibge) |
| country | (also drops state) | (also drops estado) |

The 4th dimension's name (and whether it is translated in `pt`) differs by sector:

| Dataset | eng name | pt name | pt renamed? |
|---|---|---|---|
| seeg_residuals | waste_stream | tipo_residuo | yes |
| seeg_farming | emission_pathway | via_emissao | yes |
| seeg_industry | emission_scope | recorte | **no -- stays `recorte`** |
| seeg_energy | emission_scope | recorte | **no -- stays `recorte`** |
| seeg_land | emission_scope | recorte | **no -- stays `recorte`** |

**`language = "pt"` output is never value-translated** -- only `language = "eng"` runs SEEG's
PT->EN value dictionaries (`SEEG_SECTOR_EN`, `SEEG_EMISSION_TYPE_EN`, and one
category/subcategory/dimension/activity dictionary per sector, in `R/seeg.R`). A PT call returns
the source's own Portuguese values as-is.

***

## Function Parameters

### 1. **dataset**

Selects which emission sector to download.

```r
dataset = "seeg"              # All sectors (raw_data = TRUE only)
dataset = "seeg_farming"      # Agriculture and livestock (raw_data = FALSE only)
dataset = "seeg_energy"       # Energy sector (raw_data = FALSE only)
dataset = "seeg_land"         # Land use changes (raw_data = FALSE only)
dataset = "seeg_industry"     # Industrial processes (raw_data = FALSE only)
dataset = "seeg_residuals"    # Waste and residuals (raw_data = FALSE only)
```

### 2. **raw_data**

Controls whether to download the original or the treated data. **Each dataset only works with one
value** (see above): `"seeg"` requires `TRUE`; every `seeg_*` sector requires `FALSE`.

### 3. **geo_level**

Specifies geographic aggregation level.

- `"country"`: national total
- `"state"`: state-level emissions (27 units)
- `"municipality"`: all 5,570+ municipalities

### 4. **language**

Output language for column names and, for `eng` only, translated values.

- `"pt"`: Portuguese column names, source's own Portuguese values (untranslated)
- `"eng"`: English column names and translated values

**Note on timing**: the source workbook is 250MB+; downloads may take time. A session-scoped
download+parse cache (`R/download.R`) avoids re-downloading it if you load more than one SEEG
dataset in the same session.

***

## Examples

### Example 1: All sectors combined (raw data) at the country level

```{r eval=FALSE}
# note: dataset = "seeg" only works with raw_data = TRUE
all_emissions <- load_seeg(
  dataset = "seeg",
  raw_data = TRUE,
  geo_level = "country",
  language = "eng"
)
```

### Example 2: Agricultural emissions by state

```{r eval=FALSE}
farming <- load_seeg(
  dataset = "seeg_farming",
  raw_data = FALSE,
  geo_level = "state",
  language = "eng"
)
```

### Example 3: Land use change emissions by state

```{r eval=FALSE}
land_use <- load_seeg(
  dataset = "seeg_land",
  raw_data = FALSE,
  geo_level = "state",
  language = "eng"
)
```

### Example 4: Energy emissions by municipality

```{r eval=FALSE}
energy <- load_seeg(
  dataset = "seeg_energy",
  raw_data = FALSE,
  geo_level = "municipality",
  language = "eng"
)
```

### Example 5: Industrial process emissions by state

```{r eval=FALSE}
industry <- load_seeg(
  dataset = "seeg_industry",
  raw_data = FALSE,
  geo_level = "state",
  language = "eng"
)
```

### Example 6: Waste emissions by state

```{r eval=FALSE}
residuals <- load_seeg(
  dataset = "seeg_residuals",
  raw_data = FALSE,
  geo_level = "state",
  language = "eng"
)
```

## Limitations

1. **No year selection**: entire series (1970-2024) is downloaded; there is no `time_period`
   argument.
2. **File size**: the source workbook is 250MB+; a stable connection is recommended.
3. **Missing detail**: see "Detail not included" above -- fuel type, species/crop detail,
   biome, and the land-cover transition matrix are not available in the current data.
4. **`pt` output is untranslated**: only `language = "eng"` runs SEEG's value dictionaries.

***
