---
title: "SIGMINE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIGMINE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

The Mining Geographical Information System (SIGMINE) is maintained by the Brazilian Mining National Agency (Agencia Nacional de Mineracao - ANM). It contains shapefiles of active mining licenses, leasing, blocked and protected areas and mining reserves. Data is daily updated.

The \code{load_sigmine} extracts data from a local path or directly from the website \url{https://dados.gov.br/dataset/sistema-de-informacoes-geograficas-da-mineracao-sigmine} and return a panel data at the year level since the beginning of the historical record. 

Users can choose to retrieve data from mining areas, commodities extracted, land usage and licensing steps (leasings) in each geographical unit. Only brazilian states are currently supported as geographical units, but further versions will include municipalities.

Using the function is easy enough. The example below returns all data types:

```{r setup}
library(datazoom.amazonia)

dat = load_sigmine(space_aggregation = 'state',info='all',local='web',return='edited')

```
