---
title: "SIGMINE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIGMINE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Dataset

<<<<<<< HEAD
The Mining Geographical Information System (SIGMINE) is maintained by the Brazilian Mining National Agency (Agencia Nacional de Mineracao - ANM). It contains shapefiles of active mining licenses, leasing, blocked and protected areas and mining reserves. Data is daily updated.
=======
The SIGMINE project gathers data reporting each mine in the country and their features such as year, area in square meters, company ownership, element being explored etc.
>>>>>>> master

The \code{load_sigmine} extracts data from a local path or directly from the website \url{https://dados.gov.br/dataset/sistema-de-informacoes-geograficas-da-mineracao-sigmine} and return a panel data at the year level since the beginning of the historical record. 

Users can choose to retrieve data from mining areas, commodities extracted, land usage and licensing steps (leasings) in each geographical unit. Only brazilian states are currently supported as geographical units, but further versions will include municipalities.

Using the function is easy enough. The example below returns all data types:

```{r, eval = FALSE}
library(datazoom.amazonia)
<<<<<<< HEAD
=======
load_sigmine(space_aggregation = 'state', source = NULL, language = 'eng')
```
>>>>>>> master

dat = load_sigmine(space_aggregation = 'state',info='all',local='web',return='edited')

```
