---
title: "TerraClimate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TerraClimate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

TerraClimate is a global climate and climatic water balance dataset developed by the [Climatology Lab](https://www.climatologylab.org/) at University of California, Merced. This package provides access to TerraClimate data for Brazil and the Amazon region.

This dataset provides:

- **High-resolution climate data**: Monthly climate variables at ~4km resolution globally
- **Comprehensive variables**: Temperature, precipitation, wind, radiation, soil moisture, drought indices
- **Water balance data**: Evapotranspiration, runoff, water deficit calculations
- **Temporal coverage**: Monthly data from 1958 to present (continuously updated)
- **Global coverage**: Available worldwide, subset here for Brazil
- **Satellite-derived data**: Combines satellite observations with ground station networks
- **Research quality**: Peer-reviewed, widely used in climate and ecological research

TerraClimate is essential for understanding climate variability, water availability, drought risk, agricultural potential, and climate change impacts across Brazil and the Amazon.

### Data Source and Methodology

TerraClimate data is compiled by:
- University of California Climatology Lab
- Integration of satellite and ground-based observations
- Validated against station networks
- Downscaled to ~4km global resolution
- Monthly temporal resolution with daily and subdaily estimates available

For more information, visit [TerraClimate Project](https://www.climatologylab.org/terraclimate.html).

***

## Available Climate Variables

TerraClimate provides 13 main climate and water balance variables:

| Dataset | Code | Description | Units |
|---------|------|-------------|-------|
| max_temperature | tmax | Maximum 2-m Temperature | °C |
| min_temperature | tmin | Minimum 2-m Temperature | °C |
| wind_speed | ws | Wind Speed at 10-m | m/s |
| vapor_pressure_deficit | vpd | Vapor Pressure Deficit | kPa |
| vapor_pressure | vap | 2-m Vapor Pressure | kPa |
| snow_water_equivalent | swe | Snow Water Equivalent at End of Month | mm |
| shortwave_radiation_flux | srad | Downward Shortwave Radiation Flux | W/m² |
| soil_moisture | soil | Soil Moisture at End of Month | mm |
| runoff | q | Runoff | mm |
| precipitation | ppt | Accumulated Precipitation | mm |
| potential_evaporation | pet | Reference Evapotranspiration | mm |
| climatic_water_deficit | def | Climatic Water Deficit | mm |
| water_evaporation | aet | Actual Evapotranspiration | mm |
| palmer_drought_severity_index | PDSI | Palmer Drought Severity Index | unitless |

***

## Data Format and Coverage

### Spatial Resolution

- **Resolution**: Approximately 4 km (0.04° at equator)
- **Coverage**: Global; subset available for Brazil and Legal Amazon
- **Coordinates**: WGS84 latitude/longitude

### Temporal Resolution

- **Frequency**: Monthly
- **Time span**: 1958 to present (continuously updated)
- **Data lag**: Recent months added as they become available (typically 2-3 months delay)

### Data Type

- **Format**: NetCDF files (raster/grid data)
- **Size**: Large for multi-year, multi-variable downloads
- **Access**: Downloaded from THREDDS server

***

## Function Parameters

### 1. **dataset**

Selects which climate variable to download.

```r
# Temperature and radiation
dataset = "max_temperature"         # tmax
dataset = "min_temperature"         # tmin
dataset = "shortwave_radiation_flux" # srad

# Water and moisture
dataset = "precipitation"           # ppt
dataset = "potential_evaporation"   # pet
dataset = "water_evaporation"       # aet (actual evapotranspiration)
dataset = "soil_moisture"           # soil
dataset = "runoff"                  # q

# Atmospheric variables
dataset = "wind_speed"              # ws
dataset = "vapor_pressure"          # vap
dataset = "vapor_pressure_deficit"  # vpd

# Drought and composite indices
dataset = "climatic_water_deficit"  # def
dataset = "palmer_drought_severity_index" # PDSI
dataset = "snow_water_equivalent"   # swe
```

### 2. **raw_data**

Controls data format returned.

- `TRUE`: Returns raw raster data (NetCDF, SpatRaster format)
- `FALSE`: Returns aggregated data (specific format depends on configuration)

```r
raw_data = FALSE  # logical
```

### 3. **time_period**

Specifies which year(s) to download.

**Available range**: 1958 to present (most recent months have 2-3 month lag)

```r
time_period = 2020              # single year
time_period = c(2010, 2020)     # specific years
time_period = 2010:2020         # range of years
```

### 4. **legal_amazon_only**

Restricts geographic coverage to Legal Amazon region.

- `TRUE`: Downloads only data for Legal Amazon region (much smaller files)
- `FALSE`: Downloads data for all Brazil (larger files)

```r
legal_amazon_only = TRUE  # logical
```

**Recommendation**: Use `TRUE` to significantly reduce download size for Amazon-focused research.

### 5. **language**

Output language for variable names and documentation.

- `"pt"`: Portuguese
- `"eng"`: English

```r
language = "eng"  # character string
```

***

## Important Download Considerations

 **Data Size**: TerraClimate raster data is substantial. Consider:

- **File size**: Multi-year downloads can be very large (hundreds of MB to GBs)
- **Internet**: High-speed connection recommended; THREDDS downloads can be slow
- **Storage**: Ensure sufficient disk space (multiple years of global data is large)
- **Time**: Downloads may take considerable time
- **Memory**: Raster processing requires sufficient RAM

**Recommendations**:
- Use `legal_amazon_only = TRUE` to reduce size by ~95%
- Download single or 2-3 year periods rather than decades at once
- Use high-speed internet connection
- Have at least 10-50 GB free disk space for multi-year downloads

***

## Examples

### Example 1: Precipitation Patterns

Analyze rainfall across the Amazon:

```{r eval=FALSE}
library(raster)
library(dplyr)

# Download precipitation for one year in Amazon
precip <- load_climate(
  dataset = "precipitation",
  time_period = 2020,
  legal_amazon_only = TRUE,
  language = "eng"
)

# Basic statistics
summary(precip)

# Plot precipitation
plot(precip, main = "Amazon Precipitation (2020)")

# Annual total (sum of months)
annual_precip <- sum(precip, na.rm = TRUE)
plot(annual_precip, main = "Annual Precipitation in Amazon")
```

### Example 2: Temperature Trends

Track temperature changes over time:

```{r eval=FALSE}
library(raster)

# Download max temperature for recent decade
max_temp <- load_climate(
  dataset = "max_temperature",
  time_period = 2010:2020,
  legal_amazon_only = TRUE,
  language = "eng"
)

# Average temperature over period
avg_temp <- mean(max_temp, na.rm = TRUE)
plot(avg_temp, main = "Average Maximum Temperature (2010-2020)")

# Temperature range: compare first and last year
temp_2010 <- max_temp[[1:12]]  # First year
temp_2020 <- max_temp[[(nlayers(max_temp)-11):nlayers(max_temp)]]  # Last year

avg_2010 <- mean(temp_2010, na.rm = TRUE)
avg_2020 <- mean(temp_2020, na.rm = TRUE)

# Simple warming calculation
temp_change <- avg_2020 - avg_2010
plot(temp_change, main = "Temperature Change 2010-2020")
```

### Example 3: Drought Assessment

Analyze drought conditions using water deficit data:

```{r eval=FALSE}
library(raster)

# Download climatic water deficit
water_def <- load_climate(
  dataset = "climatic_water_deficit",
  time_period = 2020,
  legal_amazon_only = TRUE,
  language = "eng"
)

# Water deficit (high values = dry conditions)
# Average deficit over year
avg_water_def <- mean(water_def, na.rm = TRUE)
plot(avg_water_def, main = "Average Climatic Water Deficit (2020)")

# Identify dry areas (high deficit)
drought_areas <- avg_water_def > quantile(avg_water_def, 0.75, na.rm = TRUE)
plot(drought_areas, main = "High Drought Risk Areas")

# Palmer Drought Severity Index for additional drought context
pdsi <- load_climate(
  dataset = "palmer_drought_severity_index",
  time_period = 2020,
  legal_amazon_only = TRUE,
  language = "eng"
)

# PDSI: negative = drought, positive = wet
plot(pdsi, main = "Palmer Drought Severity Index (2020)")
```
### Example 4: Radiation and Growing Season

Analyze solar radiation and its relationship to temperature:

```{r eval=FALSE}
library(raster)

# Download radiation and temperature
radiation <- load_climate(
  dataset = "shortwave_radiation_flux",
  time_period = 2020,
  legal_amazon_only = TRUE,
  language = "eng"
)

temp <- load_climate(
  dataset = "max_temperature",
  time_period = 2020,
  legal_amazon_only = TRUE,
  language = "eng"
)

# Annual radiation total
annual_radiation <- sum(radiation, na.rm = TRUE)
plot(annual_radiation, main = "Annual Solar Radiation")

# Average temperature
avg_temp <- mean(temp, na.rm = TRUE)
plot(avg_temp, main = "Average Maximum Temperature")

# Correlation between radiation and temperature
# (Simplified: stack and calculate)
radiation_temp <- stack(annual_radiation, avg_temp)
names(radiation_temp) <- c("Radiation", "Temperature")
```

### Example 5: Soil Moisture Dynamics

Track seasonal soil moisture patterns:

```{r eval=FALSE}
library(raster)

# Download soil moisture
soil_moisture <- load_climate(
  dataset = "soil_moisture",
  time_period = 2020,
  legal_amazon_only = TRUE,
  language = "eng"
)

# Monthly stack (12 months if 1 year)
plot(soil_moisture, main = "Monthly Soil Moisture (2020)")

# Identify wet vs. dry months
avg_soil <- mean(soil_moisture, na.rm = TRUE)

# High vs. low moisture areas
high_moisture <- soil_moisture > quantile(soil_moisture, 0.75, na.rm = TRUE)
low_moisture <- soil_moisture < quantile(soil_moisture, 0.25, na.rm = TRUE)

plot(high_moisture, main = "High Soil Moisture Areas")
plot(low_moisture, main = "Low Soil Moisture Areas")
```
***

## Data Notes

### Variable Definitions

- **Tmax/Tmin**: Monthly average maximum and minimum 2-meter air temperatures
- **Precipitation**: Accumulated monthly precipitation
- **Evapotranspiration (AET)**: Actual water loss from soil + plants through evaporation/transpiration
- **Potential Evaporation (PET)**: Theoretical maximum evapotranspiration if unlimited water
- **Runoff**: Water flowing overland/through soil to streams
- **Soil Moisture**: Water stored in root zone at month end
- **Vapor Pressure Deficit (VPD)**: Difference between saturated and actual vapor pressure (indicator of atmospheric dryness)
- **Water Deficit (DEF)**: Accumulated water stress (PET - AET)
- **PDSI**: Standardized drought index (-4 to +4 scale)

### Data Quality

- **Validation**: Validated against independent station networks
- **Uncertainty**: Varies by region; higher in data-sparse areas
- **Interpolation**: Satellite and station data combined and downscaled to 4km

- **Reliability**: Generally excellent for temperature and precipitation; water balance variables have higher uncertainty

### Important Limitations

1. **Raster data large**: Multi-year downloads can be hundreds of MB to GBs
2. **4km resolution**: Suitable for regional analysis; may miss local variation
3. **Monthly aggregation**: Daily and sub-daily variation not captured
4. **Recent data lag**: Most recent 2-3 months not yet available
5. **Interpolation uncertainty**: Some regions have lower data density
6. **Snow/ice areas**: Less accurate in high mountains or glaciated regions (not major issue for Amazon)

### Recommended Processing

Due to large file size:
- **Aggregate early**: Summarize by month/year quickly to reduce memory use
- **Legal Amazon only**: Massive size reduction if working in Amazon region
- **Subset years**: Download only years of interest rather than decades
- **Extract points**: If doing point-based analysis, extract specific coordinates to simplify data
- **Use cloud computing**: Consider cloud platforms (Google Earth Engine) for very large analyses

***

